---
title: "Untitled"
author: "Vadim Karnaukhov"
date: "19 04 2022"
output: html_document
---

# Summary of all benchmark cases

```{r}
general <- fread("mir_output/general.txt") %>% 
  mutate(pdb.id = substr(pdb.id, 1, 4))

summary <- fread("mir_output/general.txt") %>% 
  filter(chain.type %in% c("TRB", "TRA")) %>% 
  select(pdb.id, chain.type, allele.info) %>% 
  dcast(pdb.id ~ chain.type) %>% 
  mutate(TRAV = str_split_fixed(TRA, ":", 2)[,1],
         TRAJ = str_split_fixed(TRA, ":", 2)[,2],
         TRBV = str_split_fixed(TRB, ":", 2)[,1],
         TRBJ = str_split_fixed(TRB, ":", 2)[,2]) %>% 
  merge(fread("mir_output/markup.txt") %>% 
          filter(region.type == "CDR3") %>% 
          merge(fread("mir_output/general.txt") %>% 
                  select(pdb.id, chain.id, chain.type)) %>% 
          mutate(region = paste0(region.type, substr(chain.type, 3, 3))) %>% 
          select(pdb.id, region, region.sequence) %>% 
          dcast(pdb.id ~ region)) %>% 
  select(-TRA, -TRB) %>% 
  merge(fread("mir_output/markup.txt") %>% 
          filter(region.type == "PEPTIDE") %>% 
          select(pdb.id, peptide = region.sequence)) %>% 
  mutate(pdb.id = substr(pdb.id, 1, 4)) %>% 
  mutate(peptide = gsub("X", "", peptide)) %>% 
  filter(pdb.id %in% pdb_nonred)

#tibble(pdb.id = pdb_nonred) %>% fwrite("pdb_nonred.txt")
```

# TCRen

```{r}
bench.TCRen <- bench.peptides.ranks %>% 
  filter(potential == "TCRen.a.cf") %>% 
  select(pdb.id, rank = value.rank) %>% 
  merge(summary) %>% 
  mutate(method = "TCRen",
         top.1 = (rank <= 1),
         top.10 = (rank < 10),
         top.25 = (rank < 25),
         top.50 = (rank < 50),
         supported = T) %>% 
  select(pdb.id, peptide, CDR3A, CDR3B, method, rank, top.1, top.10, top.25, top.50, supported)
```


# TCRex benchmark results

```{r}
TCRex <- fread("grep -v '^#' other_tools/TCRex_results.tsv") 
TCRex.supported <- fread("other_tools/TCRex_supported_peptides.csv")

bench.TCRex <- TCRex %>% 
  select(CDR3B = CDR3_beta, peptide = epitope, score) %>% 
  distinct() %>% 
  group_by(CDR3B) %>% 
  mutate(rank = rank(-score)) %>% 
  merge(summary, all.y = T) %>% 
  merge(TCRex.supported %>% 
          mutate(supported = T), all.x = T) %>% 
  mutate(method = "TCRex",
         rank = replace_na(rank, Inf),
         top.1 = (rank == 1),
         top.10 = (rank <= 0.1 * nrow(TCRex.supported)),
         top.25 = (rank <= 0.25 * nrow(TCRex.supported)),
         top.50 = (rank <= 0.5 * nrow(TCRex.supported)),
         supported = replace_na(supported, F)) %>% 
  select(pdb.id, peptide, CDR3A, CDR3B, method, rank, top.1, top.10, top.25, top.50, supported)
```

# TITAN (on random peptide dataset)

Prepare input
```{r}
TITAN.random <- bench.peptides %>% 
  rowwise() %>% 
  filter(pdb.id %in% pdb_nonred,
         real == F) %>% 
  select(-real) %>% 
  group_by(pdb.id) %>% 
  slice_sample(n = 127)

TITAN.input.random.smiles <- TITAN.random %>% 
  select(peptide) %>% 
  rbind(summary %>% select(peptide) %>% distinct()) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(smiles = aaSMILES(peptide),
         id = row_number())

TITAN.input.random.smiles %>% 
  select(smiles, id) %>% 
  fwrite("~/git/TITAN_new/benchmark/input_epitopes_random.smi", sep = "\t", col.names=FALSE)

TITAN.input.tcr <- fread("~/git/TITAN/benchmark/input_tcr_fullseq.csv", header = T) %>% 
  mutate(id = c(1:nrow(.))) 

TITAN.input.tcr %>%   
  select(full_seq, id) %>% 
  fwrite("~/git/TITAN/benchmark/input_tcr.csv", sep = "\t", col.names=FALSE)

TITAN.input.random.test <- summary %>% 
  select(pdb.id, CDR3B) %>%   
  merge(TITAN.random, all = T) %>% 
  mutate(label = 0) %>%
  rbind(summary %>% 
          select(pdb.id, CDR3B, peptide) %>% 
          mutate(label = 1)) %>% 
  merge(TITAN.input.tcr %>% select(CDR3B = CDR3b, sequence_id = id)) %>% 
  merge(TITAN.input.random.smiles %>% select(peptide, ligand_name = id))

TITAN.input.random.test %>% 
  select(ligand_name, sequence_id, label) %>% 
  fwrite("~/git/TITAN_new/benchmark/input_test_random.csv", row.names = T)
```

Analyse output
```{r}
bench.TITAN <- fread("~/git/TITAN_new/benchmark/output_bench_random_2.csv", header = T) %>% 
  magrittr::set_colnames(c("id", "score", "label")) %>% 
  merge(TITAN.input.random.test %>% 
          mutate(id = 0:(nrow(.)-1))) %>% 
  group_by(pdb.id) %>% 
  mutate(rank = rank(-score) / n() * 100) %>% 
  filter(label == 1) %>% 
  merge(summary %>% select(pdb.id, CDR3A, CDR3B), all.y = T) %>% 
  mutate(rank = replace_na(rank, Inf)) %>% 
  mutate(method = "TITAN",
         top.1 = (rank < 1),
         top.10 = (rank < 10),
         top.25 = (rank < 25),
         top.50 = (rank < 50),
         supported = T) %>% 
  select(pdb.id, peptide, CDR3A, CDR3B, method, rank, top.1, top.10, top.25, top.50, supported)
```

# ERGO (on random peptide dataset)

Prepare input
```{r}
ERGO.input <- summary %>% 
  select(-peptide) %>% 
  merge(bench.peptides %>% 
          #group_by(pdb.id, real) %>% 
          #slice_sample(n = 200) %>% 
          select(pdb.id, peptide, real)) %>% 
  mutate(TRAV = str_split_fixed(TRAV, "\\*", 2)[,1],
         TRAJ = str_split_fixed(TRAJ, "\\*", 2)[,1],
         TRBV = str_split_fixed(TRBV, "\\*", 2)[,1],
         TRBJ = str_split_fixed(TRBJ, "\\*", 2)[,1],
         MHC = NaN) %>% 
  merge(general %>% 
          filter(chain.type == "MHCa") %>% 
          select(pdb.id, MHC.class = chain.supertype)) %>% 
  mutate(`T-Cell-Type` = ifelse(MHC.class == "MHCI", "CD8", "CD4"),
         peptide = gsub("X", "", peptide)) %>% 
  select(pdb.id, TRA = CDR3A, TRB = CDR3B, TRAV, TRAJ, TRBV, TRBJ, `T-Cell-Type`, Peptide = peptide, MHC, real)

#ERGO.input %>% select(-pdb.id , -real) %>% fwrite("~/git/ERGO-II/benchmark/input_bench_long.csv")
```

Analyse output
```{r}
bench.ERGO <- fread("~/git/ERGO-II/benchmark/output_bench_long.csv") %>% 
  merge(ERGO.input %>% select(-MHC)) %>% 
  group_by(pdb.id) %>% 
  mutate(rank = rank(-Score) / n() * 100) %>% 
  filter(real) %>% 
  mutate(method = "ERGO-II",
         top.1 = (rank < 1),
         top.10 = (rank < 10),
         top.25 = (rank < 25),
         top.50 = (rank < 50),
         supported = T) %>% 
  select(pdb.id, peptide = Peptide, CDR3A = TRA, CDR3B = TRB, method, rank, top.1, top.10, top.25, top.50, supported)
```


# ImRex (on random peptide dataset)
```{r}
ImRex.input <- bench.peptides %>% 
  rename(antigen.epitope = peptide) %>% 
  merge(summary %>% 
          select(pdb.id, cdr3 = CDR3B)) 

#fwrite(ImRex.input %>% select(antigen.epitope, cdr3) %>% distinct(), 
#       "~/git/ImRex/benchmark/ImRex_input.csv", sep = ";")
```

```{r}
bench.ImRex <- fread("~/git/ImRex/benchmark/ImRex_output.csv") %>% 
  merge(ImRex.input) %>% 
  group_by(pdb.id) %>% 
  mutate(rank = rank(-prediction_score) / n() * 100) %>% 
  filter(real) %>% 
  select(pdb.id, rank) %>% 
  merge(summary, all.y = T) %>% 
  mutate(rank = replace_na(rank, Inf)) %>% 
  mutate(method = "ImRex",
         top.1 = (rank <= 1),
         top.10 = (rank < 10),
         top.25 = (rank < 25),
         top.50 = (rank < 50),
         supported = (rank != Inf)) %>% 
  select(pdb.id, peptide, CDR3A, CDR3B, method, rank, top.1, top.10, top.25, top.50, supported)
```

# tcrdist3

Prepare tcrdist3 input for vdjdb_tcrb
```{r}
# Remove TCRs with multiple specificity and TRB from benchmark
vdjdb.trb <- fread("~/vdjdb/vdjdb-2022-03-30/vdjdb.slim.txt") %>% 
  filter(gene == "TRB") %>% 
  select(species, epitope = antigen.epitope, v_b_gene = v.segm, j_b_gene = j.segm, cdr3_b_aa = cdr3) %>% 
  distinct()

tcrdist.vdjdb <- vdjdb.trb %>% 
  filter(!grepl('\\,', v_b_gene),
         !grepl('\\,', j_b_gene)) %>% 
  group_by(cdr3_b_aa) %>% 
  mutate(n = n()) %>% 
  filter(n == 1,
         !cdr3_b_aa %in% summary$CDR3B,
         species %in% c("HomoSapiens", "MusMusculus")) %>% 
  select(-n)

#tcrdist.vdjdb %>% filter(species == "HomoSapiens") %>% fwrite("~/git/tcrdist3/benchmark/vdjdb_trb_human.csv")
#tcrdist.vdjdb %>% filter(species == "MusMusculus") %>% fwrite("~/git/tcrdist3/benchmark/vdjdb_trb_mouse.csv")
```

Prepare tcrdist3 input for bench_trb
```{r}
bench.species <- fread("mir_output/general.txt") %>% 
  mutate(pdb.id = substr(pdb.id, 1, 4)) %>% 
  select(pdb.id, species = complex.species) %>% 
  distinct()

tcrdist.bench <- summary %>% 
  merge(bench.species) %>% 
  select(species, epitope = peptide, v_b_gene = TRBV, j_b_gene = TRBJ, cdr3_b_aa = CDR3B)

#tcrdist.bench %>% filter(species == "Human") %>% fwrite("~/git/tcrdist3/benchmark/bench_trb_human.csv")
#tcrdist.bench %>% filter(species == "Mouse") %>% fwrite("~/git/tcrdist3/benchmark/bench_trb_mouse.csv")
```

Process TCRdist output
```{r}
tcrdist.human <- fread("~/git/tcrdist3/benchmark/output_dist_trv_human.csv", header = T) %>% 
  melt(id = "V1") %>% 
  set_colnames(c("id_vdjdb", "id_bench",  "tcrdist")) %>% 
  merge(fread("~/git/tcrdist3/benchmark/output_bench_trv_human.csv") %>% 
          select(id_bench = V1, cdr3_bench = cdr3_b_aa, v_bench = v_b_gene, epitope_bench = epitope)) %>% 
  merge(fread("~/git/tcrdist3/benchmark/output_vdjdb_trv_human.csv") %>% 
          select(id_vdjdb = V1, cdr3_vdjdb = cdr3_b_aa, v_vdjdb = v_b_gene, epitope_vdjdb = epitope)) %>% 
  filter(tcrdist != 0)

tcrdist.mouse <- fread("~/git/tcrdist3/benchmark/output_dist_trv_mouse.csv", header = T) %>% 
  melt(id = "V1") %>% 
  set_colnames(c("id_vdjdb", "id_bench",  "tcrdist")) %>% 
  merge(fread("~/git/tcrdist3/benchmark/output_bench_trv_mouse.csv") %>% 
          select(id_bench = V1, cdr3_bench = cdr3_b_aa, v_bench = v_b_gene, epitope_bench = epitope)) %>% 
  merge(fread("~/git/tcrdist3/benchmark/output_vdjdb_trv_mouse.csv") %>% 
          select(id_vdjdb = V1, cdr3_vdjdb = cdr3_b_aa, v_vdjdb = v_b_gene, epitope_vdjdb = epitope)) %>% 
  filter(tcrdist != 0)

tcrdist <- rbind(
  tcrdist.human %>% mutate(species = "human"),
  tcrdist.mouse %>% mutate(species = "mouse")
)
```

```{r}
tcrdist.closest <- tcrdist %>% 
  arrange(species, id_bench, tcrdist) %>% 
  group_by(species, id_bench, epitope_vdjdb) %>% 
  filter(row_number() == 1) %>% 
  merge(tcrdist.vdjdb %>% 
          mutate(species = ifelse(species == "HomoSapiens", "human", "mouse")) %>% 
          rename(epitope_vdjdb = epitope) %>% 
          group_by(species, epitope_vdjdb) %>% 
          summarise(n.vdjdb = n())) %>% 
  mutate(vdjdb_in_bench = (epitope_vdjdb %in% tcrdist.bench$epitope)) 

bench.tcrdist <- tcrdist.closest %>%
  filter(vdjdb_in_bench | n.vdjdb > 100) %>% 
  #filter(vdjdb_in_bench) %>% 
  group_by(species, id_bench) %>% 
  mutate(rank = rank(tcrdist),
         rank.perc = rank / n() * 100,
         real = epitope_bench == epitope_vdjdb) %>% 
  filter(real) %>% 
  ungroup %>% 
  select(CDR3B = cdr3_bench, peptide = epitope_bench, rank.perc, rank) %>% 
  merge(summary, all.y = T) %>% 
  mutate(rank.perc = replace_na(rank.perc, Inf),
         rank = replace_na(rank, Inf)) %>% 
  mutate(method = "tcrdist3",
         top.1 = (rank.perc <=1 | rank == 1),
         top.10 = (rank.perc <= 10 | rank == 1),
         top.25 = (rank.perc <= 25 | rank == 1),
         top.50 = (rank.perc < 50 | rank == 1),
         supported = (rank.perc != Inf)) %>% 
  select(pdb.id, peptide, CDR3A, CDR3B, method, rank = rank.perc, top.1, top.10, top.25, top.50, supported)
```



# Summarise results of all tools
```{r}
library("ggsci")

levels_methods <- c("TCRen", "TITAN", "ERGO-II", "ImRex", "TCRex", "tcrdist3")

bench.all_methods <- rbind(bench.TCRen, bench.ImRex, bench.ERGO, bench.TCRex, bench.tcrdist,
                   bench.TITAN) %>% 
  mutate(method = factor(method, levels = levels_methods)) %>% 
  replace(is.na(.), 0) 

bench.all_methods.s <- bench.all_methods %>% 
  group_by(method) %>% 
  summarise_at(.vars = c("top.1", "top.10", "top.25", "top.50", "supported"),
               .funs = "sum") %>% 
  melt(id = c("method"), variable.name = "bin", value.name = "cases") %>% 
  mutate(fraction = cases / max(cases))

p.bench.b <- bench.all_methods.s %>% 
  filter(bin == "supported") %>% 
  ggplot(aes(x = method, y = cases, fill = method)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_pubclean() +
  geom_hline(yintercept = max(bench.all_methods.s$cases), linetype = "dashed") +
  scale_y_continuous(breaks=c(50, 100, 143)) +
  ggtitle("Supported benchmark cases") +
  xlab("") + ylab("Number of cases") +
  guides(fill="none") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_fill_npg()

p.bench.c <- bench.all_methods.s %>% 
  filter(bin != "supported") %>% 
  mutate(bin = paste0(gsub("\\.", " ", bin), "%")) %>% 
  ggplot(aes(x = bin, y = fraction, fill = method)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_pubclean() +
  ggtitle("Success rate") +
  xlab("") + ylab("Success rate") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom",
        legend.margin=margin(t=-25)) +
  guides(fill = guide_legend(nrow = 1)) +
  scale_fill_npg()
```

```{r, fig.width=12, fig.height=4}
ggdraw() +
  #draw_plot(p.bench.b, x = .02, y = -.08, width = .2, height = 1.1) +
  #draw_plot(p.bench.c, x = .25, y = 0, width = .75, height = 1) +
  draw_plot(p.bench.b, x = .77, y = -.08, width = .2, height = 1.1) +
  draw_plot(p.bench.c, x = .02, y = 0, width = .75, height = 1) +
  draw_plot_label(label = c("B", "C"), size = 14,
                  x = c(0, .75), y = c(1, 1))
```

```{r}
bench.all_methods.n_vdjdb <- bench.all_methods %>%
  #filter(rank != Inf) %>% 
  #merge(vdjdb.n) %>% 
  merge(tcrdist.vdjdb %>% 
          rename(peptide = epitope) %>% 
          group_by(peptide) %>% 
          summarise(n.vdjdb = n()), all.x = T) %>% 
  mutate(n.vdjdb = replace_na(n.vdjdb, 0),
         rank = as.numeric(rank)) 

p.bench.sa <- bench.all_methods.n_vdjdb %>% 
  filter(method != "TCRex") %>% 
  ggplot(aes(x = n.vdjdb, y = rank)) +
  geom_point() +
  geom_smooth(method = "lm", scales = "free") +
  facet_wrap(~method) +
  theme_pubclean() +
  ylab("Cognate epitope rank") +
  xlab("Number of VDJdb entries for the epitope") +
  scale_x_log10()

p.bench.sa

bench.all_methods.n_vdjdb %>% 
  filter(method == "TCRex") %>% 
  mutate(group = ifelse(rank == 1, "#1",
                        ifelse(rank <= 2, "#2",
                               ifelse(supported == F, "not supported", "fail")))) %>% 
  ggplot(aes(x = group, y = n.vdjdb)) +
  geom_quasirandom() +
  geom_boxplot()
```

```{r, fig.width=5, fig.height=3}
p.bench.sb <- bench.all_methods.n_vdjdb  %>% 
  filter(method == "TCRex") %>% 
  merge(fread("other_tools/TCRex_supported_peptides.csv") %>% 
          select(peptide = Epitope, n.tcrex = `Nr. of sequences`), all.x = T) %>% 
  mutate(n.vdjdb = ifelse(!is.na(n.tcrex), n.tcrex, n.vdjdb)) %>% 
  mutate(n.vdjdb.bin = ifelse(n.vdjdb == 0, "0", 
                              ifelse(n.vdjdb <= 10, "1-10",
                                     ifelse(n.vdjdb <= 100, "10-100",
                                            ifelse(n.vdjdb <= 1000, "100-1000", ">1000")))),
         n.vdjdb.bin = factor(n.vdjdb.bin, levels = c("0", "1-10", "10-100", "100-1000", ">1000")),
         group = ifelse(rank <= 2, "Success",
                        ifelse(supported == F, "Not supported", "Fail")),
         group = factor(group, levels = c("Success", "Fail", "Not supported"))) %>% 
  group_by(n.vdjdb.bin, group, method) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(x = n.vdjdb.bin, y = n, fill = group)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
  scale_fill_manual("Prediction", values = c("#00B81F", "red", "grey40")) +
  theme_pubclean() +
  xlab("Number of specific TCR sequences") + ylab("# benchmark cases") +
  facet_wrap(~method) +
  theme(legend.position = "bottom") +
  guides(fill=guide_legend(ncol=1))

p.bench.sb
```

```{r}
library(PerformanceAnalytics)
bench.all_methods %>% 
  select(pdb.id, method, rank) %>% 
  dcast(pdb.id ~ method) %>% 
  column_to_rownames("pdb.id") %>% 
  cor()
  chart.Correlation()
  

```



```{r}
fread("other_tools/fig3_table.csv", header = T) %>% 
  column_to_rownames("V1") %>% 
  ggtexttable(theme = ttheme("blank", base_size = 8, tbody.style = )) %>%
  tab_add_hline(at.row = 1:2, row.side = "top", linewidth = 2)
```

```{r}
bench.all_methods.s %>% 
  filter(bin == "top.1")
```


```{r}
bench.all_methods %>% 
  group_by(method, supported) %>% 
  summarise(n = n()) %>% 
  mutate(frac = n / 143) %>% 
  filter(supported == F)
```
```{r, fig.width=12, fig.height=4}
ggdraw() +
  #draw_plot(p.bench.b, x = .02, y = -.08, width = .2, height = 1.1) +
  #draw_plot(p.bench.c, x = .25, y = 0, width = .75, height = 1) +
  draw_plot(p.bench.b, x = .77, y = -.06, width = .2, height = 1.06) +
  draw_plot(p.bench.c, x = .02, y = 0, width = .75, height = 1) +
  draw_plot_label(label = c("A", "B"), size = 14,
                  x = c(0, .75), y = c(1, 1))
```

```{r, fig.width=8, fig.height=4}
ggdraw() +
  #draw_plot(p.bench.b, x = .02, y = -.08, width = .2, height = 1.1) +
  #draw_plot(p.bench.c, x = .25, y = 0, width = .75, height = 1) +
  draw_plot(p.bench.sa, x = .02, y = 0, width = .67, height = 1) +
  draw_plot(p.bench.sb, x = .7, y = .3, width = .3, height = .7) #+
  #draw_plot_label(label = c("A", "B"), size = 14,
                  x = c(0, .7), y = c(1, 1))
```


# LOO

```{r}
calc_TCRen.a.cf <- function(.pdb_list = pdb_nonred, exclude = NULL, pseudocount = 1) {
  contacts.pTCR %>% 
    filter(pdb.id %in% .pdb_list,
           !pdb.id %in% exclude) %>% 
    group_by(residue.aa.from, residue.aa.to) %>% 
    summarise(count = n(), .groups = 'drop') %>% 
    merge(expand.grid(residue.aa.from = aa.levels, residue.aa.to = aa.levels), all = T) %>% 
    mutate(count = replace_na(count, 0),
           count = count + pseudocount) %>% 
           #count = count + 1/20) %>% 
    group_by(residue.aa.from) %>% 
    mutate(total.from = sum(count)) %>% 
    group_by(residue.aa.to) %>% 
    mutate(total.to = sum(count)) %>% 
    ungroup() %>% 
    mutate(total = sum(count)) %>% 
    mutate(TCRen = -log(count * total / total.to / total.from)) %>% 
    filter(residue.aa.from != "C") %>% 
    select(residue.aa.from, residue.aa.to, TCRen)
}
```


```{r}
TCRen.LOO <- pdb_nonred %>% 
  lapply(function(x) {
    calc_TCRen.a.cf(exclude = x) %>% 
    #calc_TCRen.a.cf() %>% 
      mutate(pdb.id = x)
  }) %>% 
  rbindlist() %>% 
  rename(TCRen.LOO = TCRen)
```


```{r}
TCRen.LOO.e <- bench.peptides.contacts %>% 
  merge(TCRen.LOO) %>% 
  group_by(pdb.id, peptide, real) %>% 
  summarise(value.s = sum(TCRen.LOO)) %>% 
  mutate(method = "TCRen.LOO")
```
```{r}
TCRen.LOO.ranks <- TCRen.LOO.e %>% 
  group_by(pdb.id, method) %>% 
  mutate(rank = rank(value.s) / n() * 100) %>% 
  filter(real)

TCRen.LOO.ranks %>% 
  ggplot(aes(x = method, y = rank)) +
  geom_boxplot() +
  geom_quasirandom()
```

# Random

```{r}
random.ranks <- bench.peptides.contacts %>% 
  merge(calc_TCRen.a.cf() %>% 
          mutate(random = sample(TCRen))) %>% 
  group_by(pdb.id, peptide, real) %>% 
  summarise(value.s = sum(random)) %>% 
  mutate(method = "random") %>% 
  group_by(pdb.id) %>% 
  mutate(rank = rank(value.s) / n() * 100) %>% 
  filter(real)

random.ranks %>% 
  ggplot(aes(x = method, y = rank)) +
  geom_boxplot() +
  geom_quasirandom()
```


```{r}
bench.LOO <- TCRen.LOO.ranks %>% 
  rbind(random.ranks) %>% 
  merge(summary, all.y = T) %>% 
  mutate(top.1 = (rank <= 1),
         top.10 = (rank <= 10),
         top.25 = (rank <= 25),
         top.50 = (rank < 50),
         supported = (rank != Inf)) %>% 
  select(pdb.id, peptide, CDR3A, CDR3B, method, rank, top.1, top.10, top.25, top.50, supported) 

bench.LOO %>% 
  group_by(method) %>% 
  summarise_at(.vars = c("top.1", "top.10", "top.25", "top.50", "supported"),
               .funs = "sum") %>% 
  melt(id = c("method"), variable.name = "bin", value.name = "cases") %>% 
  mutate(fraction = cases / 141) %>% 
  rbind(bench.all_methods.s) %>% 
  filter(bin != "supported") %>% 
  filter(method %in% c("TCRen.LOO", "ImRex", "ERGO-II", "TITAN", "random")) %>% 
  mutate(method = factor(method, levels = c("TCRen.LOO", "ImRex", "ERGO-II", "TITAN", "random"))) %>% 
  mutate(bin = paste0(gsub("\\.", " ", bin), "%")) %>% 
  ggplot(aes(x = bin, y = fraction, fill = method)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_pubclean() +
  ggtitle("Success rate") +
  xlab("") + ylab("Success rate") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom",
        legend.margin=margin(t=-25)) +
  guides(fill = guide_legend(nrow = 1)) +
  ggsci::scale_fill_npg()
```

```{r}
TCRen.LOO.ranks %>%
  #filter(rank != Inf) %>% 
  #merge(vdjdb.n) %>% 
  merge(tcrdist.vdjdb %>% 
          rename(peptide = epitope) %>% 
          group_by(peptide) %>% 
          summarise(n.vdjdb = n()), all.x = T) %>% 
  mutate(n.vdjdb = replace_na(n.vdjdb, 0),
         rank = as.numeric(rank)) %>% 
  filter(n.vdjdb < 3) %>% 
  filter(method != "TCRex") %>% 
  ggplot(aes(x = "TCRen.LOO", y = rank)) +
  geom_quasirandom() +
  geom_boxplot() +
  theme_pubclean() +
  ylab("Cognate epitope rank") +
  xlab("Number of VDJdb entries for the epitope")
```

```{r}
summary.nvdjdb <- summary %>% 
  merge(tcrdist.vdjdb %>% 
          rename(peptide = epitope) %>% 
          group_by(peptide) %>% 
          summarise(n.vdjdb = n()), all.x = T, .groups = 'drop') %>% 
  mutate(n.vdjdb = replace_na(n.vdjdb, 0)) 

summary.nvdjdb %>% 
  group_by(n.vdjdb) %>% 
  summarise(n = n())

pdb.unseen <- summary.nvdjdb %>% 
  filter(n.vdjdb == 0) %>% 
  .$pdb.id
```

```{r}
bench.LOO %>% 
  rbind(bench.all_methods) %>% 
  filter(pdb.id %in% pdb.unseen) %>% 
  group_by(method) %>% 
  summarise_at(.vars = c("top.1", "top.10", "top.25", "top.50", "supported"),
               .funs = "sum") %>% 
  melt(id = c("method"), variable.name = "bin", value.name = "cases") %>% 
  mutate(fraction = cases / length(pdb.unseen)) %>% 
  filter(bin != "supported") %>% 
  filter(method %in% c("TCRen.LOO", "ImRex", "ERGO-II", "TITAN", "random")) %>% 
  mutate(method = factor(method, levels = c("TCRen.LOO", "ImRex", "ERGO-II", "TITAN", "random"))) %>% 
  mutate(bin = paste0(gsub("\\.", " ", bin), "%")) %>% 
  ggplot(aes(x = bin, y = fraction, fill = method)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_pubclean() +
  ggtitle("Success rate") +
  xlab("") + ylab("Success rate") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom",
        legend.margin=margin(t=-25)) +
  guides(fill = guide_legend(nrow = 1)) +
  ggsci::scale_fill_npg()

```

```{r}
bench.LOO %>% 
  rbind(bench.all_methods) %>% 
  filter(pdb.id %in% pdb.unseen) %>% 
  group_by(method) %>% 
  summarise_at(.vars = c("top.1", "top.10", "top.25", "top.50", "supported"),
               .funs = "sum") %>% 
  melt(id = c("method"), variable.name = "bin", value.name = "cases") %>% 
  mutate(fraction = cases / length(pdb.unseen)) %>% 
  filter(bin != "supported") %>% 
  filter(method %in% c("TCRen.LOO", "ImRex", "ERGO-II", "TITAN", "random")) %>% 
  mutate(method = factor(method, levels = c("TCRen.LOO", "ImRex", "ERGO-II", "TITAN", "random"))) %>% 
  mutate(bin = paste0(gsub("\\.", " ", bin), "%")) %>% 
  ggplot(aes(x = bin, y = fraction, fill = method)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_pubclean() +
  ggtitle("Success rate") +
  xlab("") + ylab("Success rate") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom",
        legend.margin=margin(t=-25)) +
  guides(fill = guide_legend(nrow = 1)) +
  ggsci::scale_fill_npg()

```

```{r}
fread("mir_output/general.txt") %>% 
  distinct(pdb.id, complex.species) %>% 
  group_by(complex.species) %>% 
  filter(complex.species == "Mixed")
```

```{r}
fread("mir_output/general.txt") %>% 
  filter(chain.type == "PEPTIDE") %>% 
  distinct(allele.info, complex.species) %>% 
  group_by(allele.info) %>% 
  summarise(n = n()) %>% 
  arrange(-n)

fread("mir_output/general.txt") %>% 
  filter(allele.info == "ASNENMETM")
```
```{r}
fread("~/vdjdb/vdjdb-2022-03-30/vdjdb.slim.txt") %>% 
  distinct(species, antigen.epitope) %>% 
  group_by(antigen.epitope) %>% 
  summarise(n = n()) %>% 
  arrange(-n)

q <- fread("~/vdjdb/vdjdb-2022-03-30/vdjdb.slim.txt") %>% 
  filter(antigen.epitope %in% c("VVGAVGVGK", "VVVGADGVGK", "VVVGAVGVGK"))

fread("~/vdjdb/vdjdb-2022-03-30/vdjdb.slim.txt") %>% 
  mutate(mhc.type = substr(mhc.a, 1, 2)) %>% 
  distinct(mhc.type, antigen.epitope) %>% 
  group_by(antigen.epitope) %>% 
  summarise(n = n()) %>% 
  arrange(-n)

```


```{r}
summary %>% 
  merge(general %>% 
          filter(chain.type == "MHCa") %>% 
          select(pdb.id, MHC = allele.info)) %>% 
  fwrite("~/struct2020/summary_benchmark.csv")
```

```{r}
fread("~/git/pMTnet/benchmark/prediction.csv") %>% 
  ggplot(aes(x = "", y = Rank)) +
  geom_boxplot() +
  geom_quasirandom()

fread("~/git/pMTnet/benchmark/prediction.csv") %>% 
  rbind(bench.all_methods) %>% 
  merge(summary %>%
          select(pdb.id, CDR3 = CDR3B, Antigen = peptide)) %>% 
  ggplot(aes(x = "", y = Rank)) +
  geom_boxplot() +
  geom_quasirandom()
```

```{r}
bench.LOO.2 <- bench.LOO %>% 
  rbind(bench.all_methods) %>% 
  select(pdb.id, method, rank) %>% 
  rbind(fread("~/git/pMTnet/benchmark/prediction.csv") %>% 
          merge(summary %>%
                  select(pdb.id, CDR3 = CDR3B, Antigen = peptide)) %>% 
          mutate(method = "pTMnet",
                 rank = Rank * 100) %>% 
          select(pdb.id, method, rank)) 

bench.LOO.2 %>%
  mutate(unseen = (pdb.id %in% pdb.unseen)) %>% 
  ggplot(aes(x = method, y = rank, color = unseen)) +
  geom_boxplot()

bench.LOO.2 %>%
  filter(pdb.id %in% pdb.unseen) %>% 
  filter(method %in% c("TCRen.LOO", "TITAN", "ERGO-II", "random", "ImRex")) %>% 
  ggplot(aes(x = method, y = rank)) +
  geom_violin()

bench.LOO.2 %>%
  filter(pdb.id %in% pdb.unseen,
         rank != Inf) %>% 
  filter(method %in% c("TCRen.LOO", "TITAN", "ERGO-II", "random", "ImRex")) %>% 
  ggplot(aes(x = method, y = rank)) +
  geom_boxplot()
```
```{r}
bench.LOO.2 %>% 
  filter(method == "ImRex")

expand.grid(method = c("TCRen.LOO", "TITAN", "ERGO-II", "random", "ImRex", "TCRen"),
            thr = c(1, 5, 10, 20, 25, 30, 40, 50, 100)) %>% 
  merge(bench.LOO.2 %>% 
          filter(rank != Inf) %>% 
          filter(pdb.id %in% pdb.unseen) %>% 
          group_by(method) %>% 
          mutate(total = n())) %>% 
  group_by(method, thr) %>% 
  summarise(n = sum(rank <= thr),
            fraction = n / total) %>% 
  distinct() %>% 
  ggplot(aes(x = thr, y = fraction, group = method, color = method)) +
  geom_line()

bench.LOO.2 %>% 
  merge()
  ggplot(aes(x = rank, color = method)) + 
  stat_bin(data=subset(bench.LOO.2, method=="TCRen.LOO"),aes(y=cumsum(..count..)),geom="step")
```


```{r}
bench.peptides.aaindex <- bench.peptides.contacts %>% 
  merge(aaindex.pot.sel %>% 
          filter(potential %in% c("MIYS", "KESO")) %>% 
          mutate(potential = ifelse(potential == "MIYS", "MJ", "Keskin"))) %>% 
  group_by(pdb.id, peptide, real, potential) %>% 
  summarise(value.s = sum(value))
```

```{r}
bench.peptides.aaindex.ranks <- bench.peptides.aaindex %>% 
  group_by(pdb.id, potential) %>% 
  mutate(value.rank = rank(value.s) / n() * 100) %>% 
  filter(real)

bench_MJ <- bench.peptides.aaindex.ranks %>% 
  select(pdb.id, method = potential, rank = value.rank) %>% 
  merge(summary) %>%
  mutate(top.1 = (rank <=1),
         top.10 = (rank <= 10),
         top.25 = (rank <= 25),
         top.50 = (rank < 50),
         supported = T) %>% 
  select(pdb.id, peptide, CDR3A, CDR3B, method, rank, top.1, top.10, top.25, top.50, supported)

```

```{r}
methods <- c("TCRen.LOO", "ERGO-II", "TITAN", "random", "MJ", "Keskin", "TCRen full")
bins <- c("top.10", "top.25")

bench.LOO %>% 
  rbind(bench.all_methods) %>% 
  mutate(method = ifelse(method == "TCRen", "TCRen full", method)) %>% 
  rbind(bench_MJ) %>% 
  filter(pdb.id %in% pdb.unseen) %>% 
  group_by(method) %>% 
  summarise_at(.vars = c("top.1", "top.10", "top.25", "top.50", "supported"),
               .funs = "sum") %>% 
  melt(id = c("method"), variable.name = "bin", value.name = "cases") %>% 
  filter(bin %in% bins) %>% 
  mutate(fraction = cases / length(pdb.unseen)) %>% 
  filter(bin != "supported") %>% 
  filter(method %in% methods) %>% 
  mutate(method = factor(method, methods)) %>% 
  mutate(bin = paste0(gsub("\\.", " ", bin), "%")) %>% 
  ggplot(aes(x = bin, y = fraction, fill = method)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_pubclean() +
  ggtitle("Success rate") +
  xlab("") + ylab("Success rate") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="right",
        legend.margin=margin(t=-25)) +
  guides(fill = guide_legend(nrow = 1)) +
  ggsci::scale_fill_npg() +
  theme(aspect.ratio = .4) +
  guides(fill=guide_legend(ncol=1))
```

```{r}
q <- fread("~/vdjdb/vdjdb-2022-03-30/vdjdb.slim.txt") %>% 
  filter(vdjdb.score != 0) %>% 
  group_by(antigen.epitope, species, antigen.species) %>% 
  summarise(n = n()) %>% 
  merge(summary %>% 
          select(pdb.id, antigen.epitope = peptide) %>% 
          group_by(antigen.epitope) %>% 
          summarise(n.pdb = n()), all = T) %>% 
  merge(fread("~/shared-with-me/struct2020/benchmark/final_version/other_tools/TCRex_supported_peptides.csv") %>% 
          rename(antigen.epitope = Epitope), all = T)  %>% 
  arrange(-n) 

```

```{r}
fread("~/vdjdb/vdjdb-2022-03-30/vdjdb.slim.txt") %>% 
  filter(antigen.epitope == "KLGGALQAK") %>% 
  arrange(-vdjdb.score)
```
```{r}
fread("~/vdjdb/vdjdb-2022-03-30/vdjdb.slim.txt") %>% 
  filter
```

```{r}
fread("~/vdjdb/vdjdb-2022-03-30/vdjdb.txt") %>% 
  .$web.method.seq %>% table()

fread("~/vdjdb/vdjdb-2022-03-30/vdjdb.slim.txt")$reference.id %>% unique()

```

```{r}
trab_pair <- function(df) {
  df %>% 
    filter(gene == "TRA") %>% 
    select(complex.id, TRA_CDR3 = cdr3, TRAV = v.segm, TRAJ = j.segm, mhc.a, antigen.epitope) %>% 
    merge(df %>% 
            filter(gene == "TRB") %>% 
            select(complex.id, TRB_CDR3 = cdr3, TRBV = v.segm, TRBJ = j.segm, mhc.a, antigen.epitope, reference.id)) %>% 
    distinct(TRA_CDR3, TRB_CDR3, .keep_all = T)
}

fread("~/vdjdb/vdjdb-2022-03-30/vdjdb.txt") %>% 
  filter(reference.id == "PMID:27198675") %>% 
  trab_pair()

# Studies with crystals
fread("~/vdjdb/vdjdb-2022-03-30/vdjdb.txt") %>% 
  filter(reference.id == "PMID:32461371") %>% 
  trab_pair()

fread("~/vdjdb/vdjdb-2022-03-30/vdjdb.txt") %>% 
  filter(reference.id == "PMID:32518267") %>% 
  trab_pair()


```

