---
title: "Running TCRen pipeline"
author: "Vadim Karnaukhov"
output: html_document
---

For its work the script requires file "mir-1.0-SNAPSHOT.jar", located in the same folder

# Paths for input files 

```{r}
structures <- "../example/input_structures/"  # name of directory with input structures
candidates <- "../example/candidate_epitopes.txt"  # name of file with candidate epitopes
potential <- "TCRen_potential.csv"  # name of file with an energy potential
out <- "../example/output_TCRen/"  # name of directory for TCRen output
memory <- "1G"  # memory allocation
```


# Load required packages

```{r}
library(data.table)
library(tidyverse)
library(magrittr, warn.conflicts = FALSE)
library(optparse)
library(stringr)
```


# Running MIR software devoted for annotation of TCR-peptide-MHC structures 

Function for running MIR software 
```{r}
run_mir <- function(input_str = file.path(structures, dir(structures)),
                    mir_path = file.path("mir-1.0-SNAPSHOT.jar"),
                    output_dir = file.path(out, "structures_annotation"),
                    arg, #"annotate-structures", "compute-pdb-geom", "compute-pdb-contacts"
                    print_log = T) {
  .pdb_list <- input_str %>% paste(collapse = " ")
  dir.create(file.path(out), showWarnings = FALSE)
  cmd <- str_glue("java -Xmx{memory} -cp {mir_path} com.milaboratory.mir.scripts.Examples {arg} -I {.pdb_list} -O ./{output_dir}/")
  code <- system(cmd, ignore.stdout = !print_log, ignore.stderr = !print_log)
  if(code != 0) {
    stop(str_glue("Failed to execute '{cmd}'"))
  }
}
system("java -version")
```

Running MIR software for annotation of structures and extraction of contacts 
```{r}
run_mir(arg = "annotate-structures")
run_mir(arg = "compute-pdb-contacts")
```


# Extraction of TCR-peptide contacts

Create a dataframe with all the contacts in the TCR-peptide-MHC structures
```{r}
# Consider contacts between two residues if any two atoms of them are located at distance <5 Angstrom
contacts <- fread(file.path(out, "structures_annotation", "atomdist.txt")) %>%
  filter(dist <= 5, chain.id.from != chain.id.to) %>%
  select(-atom.from, -atom.to, -dist) %>%
  unique

# Symmetrize the dataframe with the contacts   
contacts.full <- rbind(
  contacts,
  contacts %>% 
    rename(chain.id.from = chain.id.to,
           chain.id.to = chain.id.from,
           residue.index.from = residue.index.to,
           residue.index.to = residue.index.from)
)
```

Prepare annotation files 
```{r}
# Dataframe with assignment of chains
general_resmarkup <- fread(file.path(out, "structures_annotation", "general.txt")) %>% 
  merge(fread(file.path(out, "structures_annotation", "resmarkup.txt"), blank.lines.skip = T)) %>% 
  group_by(pdb.id, chain.id, region.type) %>% 
  mutate(region.start = min(residue.index)) %>% 
  ungroup 

# Dataframe with the sequence of all regions of TCR-peptide-MHC complex  
markup <- fread(file.path(out, "structures_annotation", "markup.txt"))

# Dataframe with the lengths of the peptides in the input structures
peptides_pdb <- markup %>% 
  filter(region.type == "PEPTIDE") %>% 
  select(pdb.id, peptide = region.sequence) %>% 
  mutate(pep.len = nchar(peptide)) %>% 
  ungroup()
```

Create a dataframe with TCR-peptide contacts in the input structures
```{r}
# Filter out TCR-peptide contacts
contact.map <- contacts.full %>% 
  merge.data.frame(general_resmarkup %>% 
          select(pdb.id, chain.type.from = chain.type, chain.id.from = chain.id,
                 residue.index.from = residue.index, residue.aa.from = residue.aa,
                 region.type.from = region.type, region.start.from = region.start)) %>% 
  merge.data.frame(general_resmarkup %>% 
          select(pdb.id, chain.type.to = chain.type, chain.id.to = chain.id, 
                 residue.index.to = residue.index, residue.aa.to = residue.aa,
                 region.type.to = region.type, region.start.to = region.start)) %>% 
  mutate(pos.from = residue.index.from - region.start.from,
         pos.to = residue.index.to - region.start.to) %>% 
  filter(chain.type.from %in% c("TRA", "TRB"),
         chain.type.to == "PEPTIDE") %>% 
  select(pdb.id, chain.type.from, region.type.from, 
         pos.from, pos.to, residue.aa.from, residue.aa.to)
```


# Calculation of TCRen scores for candidate epitopes

Ensure that the length of candidate epitopes is the same as in the input structures
```{r}
# For each input structure we consider only those candidate epitopes which have the same length
peptide_mut <- fread(file.path(candidates)) %>%
  mutate(pep.len = nchar(peptide)) %>% 
  merge(peptides_pdb %>% select(pdb.id, pep.len)) %>% 
  select(-pep.len)

pep_len_max <- nchar(peptide_mut$peptide) %>% max()
```

Contact maps for candidate epitopes
```{r}
# We consider the same contacts (in terms of residue numbers) as in the input structure but with substituted amino acids for peptide residues 
contacts.mut.pep <- peptide_mut %>% 
  separate(peptide, into = as.character(c(0:(pep_len_max-1))), sep = 1:pep_len_max, remove = F) %>% 
  as.data.table() %>% 
  melt(id = c("peptide", "pdb.id"), variable.name = "pos.to", value.name = "residue.aa.to") %>%
  mutate(pos.to = as.integer(as.character(pos.to))) %>% 
  merge(contact.map %>% 
          select(-residue.aa.to), allow.cartesian=TRUE) 
```

Read a file with TCRen energy potential
```{r}
# Other potentials can also be used
potential <- fread(file.path(potential)) %>% 
  melt(id = c("residue.aa.from", "residue.aa.to"), variable.name = "potential")
```

Calculate TCRen scores for candidate epitopes
```{r}
energy.mut.pep <- contacts.mut.pep %>% 
  merge(potential, by = c("residue.aa.from", "residue.aa.to")) %>% 
  group_by(pdb.id, peptide, potential) %>% 
  summarise(score = sum(value), .groups = "keep") %>% 
  arrange(pdb.id, score) %>%
  select(complex.id = pdb.id, peptide, potential, score)
```

Write an output file with candidate epitopes ranked according TCRen score
```{r}
fwrite(energy.mut.pep,
       file.path(out, "candidate_epitopes_TCRen.csv"))
```
