---
title: "Untitled"
output: html_document
---

For its work the script requires (in the same folder):
1) mir-1.0-SNAPSHOT.jar
2) PDB(s) structures for in silico peptide screening 
3) list of PDB file names, each in a separate line, no header
4) (optional) df with peptides for peptide screening (1st col: PDB name, 2nd col: peptide)
5) (optional) df with CDR3s for CDR3 screening (1st col: PDB name, 2nd col: chain (TRA or TRB), 3rd col: CDR3 seq)

```{r}
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(magrittr)
```
```{r}
library.path <- .libPaths()
library("data.table", lib.loc = library.path)
library("dplyr", lib.loc = library.path)
library("tidyr", lib.loc = library.path)
library("ggplot2", lib.loc = library.path)
library("magrittr", lib.loc = library.path)
library("optparse", lib.loc = library.path)
```


# Input data

```{r}
library("optparse")
 
option_list = list(
  make_option(c("-s", "--structures"), type="character", default="input_structures", 
              help="name of directory with input structures [default= %default]",
              metavar="character"),
  make_option(c("-c", "--candidates"), type="character", default="candidate_epitopes.tsv", 
              help="name of file with candidate epitopes [default= %default]", 
              metavar="character"),
  make_option(c("-tcren", "--tcren_files"), type="character", default="TCRen_files", 
              help="name of directory with TCRen.csv and mir-1.0-SNAPSHOT.jar files [default= %default]",
              metavar="character"),
  make_option(c("-o", "--out"), type="character", default="output_TCRen", 
              help="name of directory with TCRen output [default= %default]",
              metavar="character"),
  make_option(c("-m", "--memory"), type="character", default="20G", 
              help="memory allocation [default= %default]",
              metavar="character")
); 
 
opt_parser = OptionParser(option_list=option_list);
opt = parse_args(opt_parser);
```

```{r}
opt <- list()
opt$structures <- "input_structures"
opt$candidates <- "candidate_epitopes.tsv"
opt$out <- "output_TCRen"
opt$tcren_files <- "TCRen_files"
opt$memory <- "20G"
```

```{r}
opt <- list()
opt$structures <- "/home/vk/struct2020/stronen/TCRmodel/"
opt$candidates <- "/home/vk/struct2020/stronen/stronen_candidate_peptides.csv"
opt$out <- "/home/vk/struct2020/stronen/output_TCRen"
opt$tcren_files <- "TCRen_files"
opt$memory <- "20G"
```

```{r}
setwd(file.path(opt$tcren_files))
```


#Annotation of structures and contacts by MIR
```{r}
MEMORY <- opt$memory

run_mir <- function(input_str = file.path(opt$structures, dir(opt$structures)),
                    #input_str = paste0(opt$structures, "/", dir(opt$structures)),
                    mir_path = file.path(opt$tcren_files, "mir-1.0-SNAPSHOT.jar"),
                    output_dir = file.path(opt$out, "structures_annotation"),
                    #output_dir = paste0(opt$out, "/", "structures_annotation"),
                    arg, #"annotate-structures", "compute-pdb-geom", "compute-pdb-contacts"
                    print_log = T) {
  .pdb_list <- input_str %>% paste(collapse = " ")
  dir.create(file.path(opt$out), showWarnings = FALSE)
  dir.create(file.path(output_dir), showWarnings = FALSE)
  cmd <- str_glue("java -Xmx{MEMORY} -cp {mir_path} com.milaboratory.mir.scripts.Examples {arg} -I {.pdb_list} -O {output_dir}/")
  code <- system(cmd,
                 ignore.stdout = !print_log, 
                 ignore.stderr = !print_log)
  if(code != 0) {
    stop(str_glue("Failed to execute '{cmd}'"))
  }
}
system("java -version")
```

```{r}
run_mir(arg = "annotate-structures")
run_mir(arg = "compute-pdb-contacts")
```

# Processing of MIR output
```{r}
contacts <- fread(file.path(opt$out, "structures_annotation", "atomdist.txt")) %>%
  filter(dist <= 5, chain.id.from != chain.id.to) %>%
  select(-atom.from, -atom.to, -dist) %>%
  unique
  
contacts.full <- rbind(
  contacts,
  contacts %>% 
    rename(chain.id.from = chain.id.to,
           chain.id.to = chain.id.from,
           residue.index.from = residue.index.to,
           residue.index.to = residue.index.from)
)

general_resmarkup <- fread(file.path(opt$out, "structures_annotation", "general.txt")) %>% 
  merge(fread(file.path(opt$out, "structures_annotation", "resmarkup.txt"),
        blank.lines.skip = T)) %>% 
  group_by(pdb.id, chain.id, region.type) %>% 
  mutate(region.start = min(residue.index)) %>% 
  ungroup 
  
markup <- fread(file.path(opt$out, "structures_annotation", "markup.txt"))

peptides_pdb <- markup %>% 
  filter(region.type == "PEPTIDE") %>% 
  select(pdb.id, peptide = region.sequence) %>% 
  mutate(pep.len = nchar(peptide))
```

```{r}
contact.map <- contacts.full %>% 
  merge.data.frame(general_resmarkup %>% 
          select(pdb.id, chain.type.from = chain.type, chain.id.from = chain.id,
                 residue.index.from = residue.index, residue.aa.from = residue.aa,
                 region.type.from = region.type, region.start.from = region.start)) %>% 
  merge.data.frame(general_resmarkup %>% 
          select(pdb.id, chain.type.to = chain.type, chain.id.to = chain.id, 
                 residue.index.to = residue.index, residue.aa.to = residue.aa,
                 region.type.to = region.type, region.start.to = region.start)) %>% 
  mutate(pos.from = residue.index.from - region.start.from,
         pos.to = residue.index.to - region.start.to) %>% 
  filter(chain.type.from %in% c("TRA", "TRB"),
         chain.type.to == "PEPTIDE") %>% 
  select(pdb.id, chain.type.from, region.type.from, 
         pos.from, pos.to, residue.aa.from, residue.aa.to)
```


# Generate contact maps for mutated complexes

## peptide mutation
```{r}
peptide_mut <- fread(file.path(opt$candidates)) %>% 
  merge(peptides_pdb %>% select(pdb.id, pep.len)) %>% 
  filter(nchar(peptide) == pep.len) %>% 
  select(-pep.len)

pep_len_max <- nchar(peptide_mut$peptide) %>% max()

contacts.mut.pep <- peptide_mut %>% 
  separate(peptide, into = as.character(c(0:(pep_len_max-1))), sep = 1:pep_len_max, remove = F) %>% 
  melt(id = c("pdb.id", "peptide"), variable.name = "pos.to", value.name = "residue.aa.to") %>%
  mutate(pos.to = as.integer(as.character(pos.to))) %>% 
  filter(residue.aa.to != "") %>% 
  merge(contact.map %>% 
          select(-residue.aa.to), allow.cartesian=TRUE) 
```

#Potential
```{r}
potential <- fread(file.path(opt$tcren_files, "TCRen.csv")) %>% 
  melt(id = c("residue.aa.from", "residue.aa.to"), variable.name = "potential")
```

# Rank peptides
```{r}
energy.mut.pep <- contacts.mut.pep %>% 
  merge(potential) %>% 
  group_by(pdb.id, peptide, potential) %>% 
  summarise(value.s = sum(value)) %>% 
  arrange(pdb.id, value.s)

fwrite(energy.mut.pep,
       file.path(opt$out, "candidate_epitopes_ranked_TCRen.csv"))
```


