---
title: "Untitled"
author: "Vadim Karnaukhov"
date: "17 06 2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(forcats)
library(ggrepel)
library(broom)
library(rstatix)
library(ggpubr)
library(ggbeeswarm)
```


```{r}
aa.levels <- c("L","F","I","M","V","W","Y","C","H","A","G","P","T","S","Q","N","D","E","R","K")
```

# All PDB structures of TCR-peptide-MHC complexes
```{r}
pdb_dir <- "PDB_structures/"
pdb_list <- dir(pdb_dir)
```


#Annotation of structures and contacts by MIR
```{r}
MEMORY <- "20G"

run_mir <- function(pdb_dir,
                     mir_path = "mir-1.0-SNAPSHOT.jar", 
                     .output_dir = "mir_output",
                     arg, #"annotate-structures", "compute-pdb-geom", "compute-pdb-contacts"
                     print_log = T) {
  pdb_list <- dir(pdb_dir)
  .pdb_list <- paste0(pdb_dir, pdb_list) %>% paste(collapse = " ")
  output_dir <- paste0(pdb_dir, "/../", .output_dir)
  dir.create(output_dir, showWarnings = FALSE)
  cmd <- str_glue("java -Xmx{MEMORY} -cp {mir_path} com.milaboratory.mir.scripts.Examples {arg} -I {.pdb_list} -O ./{output_dir}/")
  code <- system(cmd,
                 ignore.stdout = !print_log, 
                 ignore.stderr = !print_log)
  if(code != 0) {
    stop(str_glue("Failed to execute '{cmd}'"))
  }
}
system("java -version")
```

```{r}
run_mir(pdb_dir = "PDB_structures/", arg = "annotate-structures")
run_mir(pdb_dir = "PDB_structures/", arg = "compute-pdb-contacts")
```

# Extract all TCR:peptide contacts from PDB structures

```{r}
contacts.1 <- fread("mir_output/atomdist.txt") %>% 
  mutate(pdb.id = substr(pdb.id, 1, 4)) %>%
  filter(dist <= 5, chain.id.from != chain.id.to) %>%
  select(-atom.from, -atom.to, -dist) %>%
  unique

contacts.2 <- contacts.1 %>%
  select(pdb.id, chain.id.from = chain.id.to, chain.id.to = chain.id.from,
         residue.index.from = residue.index.to, residue.index.to = residue.index.from)
  
contacts <- contacts.1 %>%
  rbind(contacts.2)
```

```{r}
resmarkup <- fread("mir_output/resmarkup.txt", blank.lines.skip = T) %>% 
  mutate(pdb.id = substr(pdb.id, 1, 4)) 
  

general <- fread("mir_output/general.txt") %>% 
  mutate(pdb.id = substr(pdb.id, 1, 4))

resmarkup.a <- resmarkup %>%
  merge(general)

contacts.a <- contacts %>%
  merge(resmarkup.a %>% select(pdb.id,
                               chain.id.from = chain.id,
                               chain.type.from = chain.type,
                               chain.supertype.from = chain.supertype,
                               complex.species,
                               region.type.from = region.type,
                               residue.index.from = residue.index,
                               residue.aa.from = residue.aa)) %>%
  merge(resmarkup.a %>% select(pdb.id,
                               chain.id.to = chain.id,
                               chain.type.to = chain.type,
                               chain.supertype.to = chain.supertype,
                               region.type.to = region.type,
                               residue.index.to = residue.index,
                               residue.aa.to = residue.aa)
        #, by = c("pdb.id", "chain.id.to", "residue.index.to") 
  )
```

```{r}
contacts.pTCR <- contacts.a %>% 
  filter(chain.supertype.from == "TRAB",
         chain.supertype.to == "PEPTIDE")
```

# Cluster PDB structures 

## Based on sequence similarity
```{r}
markup <- fread("mir_output/markup.txt") %>% 
  mutate(pdb.id = substr(pdb.id, 1, 4))

markup.seq <- markup %>% 
  merge(general) %>% 
  filter(chain.supertype == "PEPTIDE" | 
           chain.supertype == "TRAB" & region.type == "CDR3") %>% 
  mutate(region.type = ifelse(chain.type == "TRA", "cdr3a",
                              ifelse(chain.type == "TRB", "cdr3b", region.type %>% tolower()))) %>% 
  select(pdb.id, region.type, region.sequence) %>% 
  dcast(pdb.id ~ region.type) %>% 
  mutate(cdr3a.len = nchar(cdr3a),
         cdr3b.len = nchar(cdr3b),
         peptide.len = nchar(peptide))

dist.seq <- expand.grid(pdb.id.1 = markup.seq$pdb.id, pdb.id.2 = markup.seq$pdb.id) %>%
  merge(markup.seq %>% select(pdb.id.1 = pdb.id, cdr3a.1 = cdr3a, cdr3b.1 = cdr3b, peptide.1 = peptide)) %>%
  merge(markup.seq %>% select(pdb.id.2 = pdb.id, cdr3a.2 = cdr3a, cdr3b.2 = cdr3b, peptide.2 = peptide)) %>%
  mutate(dist.cdr3a = stringdist(cdr3a.1, cdr3a.2, method = 'dl'),
         dist.cdr3b = stringdist(cdr3b.1, cdr3b.2, method = 'dl'),
         dist.peptide = stringdist(peptide.1, peptide.2, method = 'dl')) %>% 
  rowwise() %>% 
  mutate(dist.sum = sum(dist.cdr3a + dist.cdr3b + dist.peptide))
```

```{r}
clusters.seq <- dist.seq %>% 
  select(pdb.id.1, pdb.id.2, dist.sum) %>%
  spread(pdb.id.2, dist.sum) %>%
  column_to_rownames("pdb.id.1") %>%
  as.dist() %>%
  hclust() %>%
  cutree(h = 6)
```

## Based on interface contact similarity

```{r}
contacts.pdb <- contacts.pTCR %>% 
  group_by(pdb.id, residue.aa.from, residue.aa.to) %>% 
  summarise(n = n()) %>% 
  merge(expand.grid(pdb.id = unique(contacts.pTCR$pdb.id), 
                    residue.aa.to = aa.levels,
                    residue.aa.from = aa.levels), 
        all.y = T) %>% 
  mutate(n = replace_na(n, 0),
         contact = paste0(residue.aa.from, residue.aa.to)) 

contacts.pdb.s <- contacts.pTCR %>% 
  group_by(pdb.id) %>% 
  summarise(n = n()) %>% 
  arrange(-n)

dist.cont <- contacts.pdb %>% 
  select(pdb.id, contact, n) %>% 
  dcast(pdb.id ~ contact) %>% 
  column_to_rownames("pdb.id") %>% 
  dist(method = "manhattan") %>% 
  as.matrix() %>% 
  melt(value.name = "dist.manh") %>% 
  rename(pdb.id.1 = Var1, pdb.id.2 = Var2) %>% 
  merge(contacts.pdb.s %>% select(pdb.id.1 = pdb.id, n.1 = n)) %>% 
  merge(contacts.pdb.s %>% select(pdb.id.2 = pdb.id, n.2 = n)) %>% 
  mutate(dist = dist.manh / (n.1 + n.2))
```

```{r}
clusters.cont <- dist.cont %>% 
  select(pdb.id.1, pdb.id.2, dist) %>% 
  dcast(pdb.id.1 ~ pdb.id.2) %>% 
  column_to_rownames("pdb.id.1") %>% 
  as.dist(.) %>% 
  hclust() %>% 
  cutree(., h = 0.7)
```

## Combine sequnce- and contact-based clustering results

```{r}
vector_to_df <- function(vector, col_name1, col_name2) {
  vector %>% 
    as.data.frame() %>% 
    rownames_to_column() %>% 
    set_colnames(c(col_name1, col_name2))
}

clusters.merged <- merge(
  vector_to_df(clusters.seq, "pdb.id", "cluster.seq"),
  vector_to_df(clusters.cont, "pdb.id", "cluster.cont")
  ) %>% 
  #group_by(cluster.seq) %>% 
  #mutate(n = n_distinct(cluster.cont)) %>% 
  #arrange(cluster.seq, cluster.cont) %>% 
  group_by(cluster.seq, cluster.cont) %>% 
  arrange(cluster.seq, cluster.cont, pdb.id)

pdb_nonred  <- clusters.merged %>% 
  slice(1) %>% 
  .$pdb.id
```

```{r}
table.s1 <- markup.seq %>% 
  filter(pdb.id %in% pdb_nonred) %>% 
  merge(fread("mir_output/general.txt") %>% 
          mutate(pdb.id = substr(pdb.id, 1, 4)) %>% 
          filter(chain.type == "MHCb") %>% 
          select(pdb.id, complex.species, mhc.class = chain.supertype)) %>% 
  select(pdb.id, complex.species, mhc.class, peptide, cdr3a, cdr3b) %>% 
  mutate(pdb.id = toupper(pdb.id)) %>% 
  magrittr::set_colnames(c("PDB ID", "Species", "MHC class", "peptide", "CDR3a", "CDR3b"))

table.s2 <- fread("mir_output/general.txt") %>% 
  mutate(pdb.id = toupper(substr(pdb.id, 1, 4))) %>% 
  select(pdb.id) %>% 
  distinct() %>% 
  arrange(pdb.id)
```


# TCRen computation

## Summarise counts
```{r}
#Compute asymmetric counts
counts.asym <- contacts.pTCR %>% 
  filter(pdb.id %in% pdb_nonred) %>% 
  group_by(residue.aa.from, residue.aa.to) %>% 
  summarise(count = n()) %>% 
  merge(expand.grid(residue.aa.from = aa.levels,
                    residue.aa.to = aa.levels),
        all = T) %>% 
  mutate(count = replace_na(count, 0))

#Compute symmetric counts
counts.sym <- counts.asym %>% 
  rbind(counts.asym %>% 
          select(residue.aa.from = residue.aa.to,
                 residue.aa.to = residue.aa.from,
                 count)) %>% 
  group_by(residue.aa.from, residue.aa.to) %>% 
  summarise(count = sum(count)) %>% 
  ungroup()

# Add pseudocounts

counts.asym.p <- counts.asym %>% 
  mutate(count = count + 1/20,
         freq.obs = count / sum(count))

counts.sym.p <- counts.sym %>% 
  mutate(count = count + 1/20,
         freq.obs = count / sum(count))
```

## "Contact fraction" reference state

```{r}
rs.cf <- rbind(
  counts.asym.p %>% mutate(potential = "TCRen.a.cf"),
  counts.sym.p %>% mutate(potential = "TCRen.s.cf")
) %>% 
  group_by(potential, residue.aa.from) %>% 
  mutate(total.from = sum(count)) %>% 
  group_by(potential, residue.aa.to) %>% 
  mutate(total.to = sum(count)) %>% 
  filter(residue.aa.from == residue.aa.to) %>% 
  group_by(potential) %>% 
  mutate(freq.from = total.from / sum(total.from),
         freq.to = total.to / sum(total.to)) %>% 
  ungroup %>% 
  select(residue.aa = residue.aa.from, potential, freq.from, freq.to)
```


## "Mole-fraction" reference state

```{r}
library(protr)

rs.mf.peptide <- markup.seq %>% 
  filter(pdb.id %in% pdb_nonred) %>% 
  .$peptide %>% 
  paste0(collapse = "") %>%
  gsub(pattern = "[XU]", replacement = "") %>% #delete non-standard amino acids
  extractAAC()
  
rs.mf.tcr <- markup.seq %>% 
  filter(pdb.id %in% pdb_nonred) %>%
  mutate(cdr3.trim = paste0(substr(cdr3a, 5, nchar(cdr3a) - 4), substr(cdr3b, 5, nchar(cdr3b) - 4))) %>% 
  .$cdr3.trim %>% 
  paste0(collapse = "") %>%
  gsub(pattern = "[XU]", replacement = "") %>% #delete non-standard amino acids
  extractAAC()

rs.mf.sym <- markup.seq %>% 
  filter(pdb.id %in% pdb_nonred) %>%
  mutate(interface = paste0(peptide, 
                            substr(cdr3a, 5, nchar(cdr3a) - 4), substr(cdr3b, 5, nchar(cdr3b) - 4))) %>% 
  .$interface %>% 
  paste0(collapse = "") %>%
  gsub(pattern = "[XU]", replacement = "") %>% #delete non-standard amino acids
  extractAAC()

rs.mf <- merge(
  vector_to_df(rs.mf.tcr, "residue.aa", "freq.from"),
  vector_to_df(rs.mf.peptide, "residue.aa", "freq.to")
) %>% 
  mutate(potential = "TCRen.a.mf") %>% 
  rbind(vector_to_df(rs.mf.sym, "residue.aa", "freq.from") %>% 
          merge(vector_to_df(rs.mf.sym, "residue.aa", "freq.to")) %>% 
          mutate(potential = "TCRen.s.mf"))
```

## "Normalized mole-fraction" reference state

```{r}
contacts.pTCR.a <- contacts.pTCR %>% 
  filter(pdb.id %in% pdb_nonred) %>% 
  merge(markup %>% select(pdb.id, chain.id.from = chain.id, region.type.from = region.type,
                          region.start.from = region.start)) %>% 
  mutate(pos.from = residue.index.from - region.start.from) 

mhc_classes <- general %>% 
  filter(chain.supertype %in% c("MHCI", "MHCII")) %>% 
  select(pdb.id, mhc.class = chain.supertype) %>% 
  distinct()

cdr3_seq <- markup %>% 
  merge(general) %>% 
  filter(region.type == "CDR3") %>% 
  select(pdb.id, chain.type.from = chain.type, region.sequence) %>% 
  mutate(cdr3.len = nchar(region.sequence))

rs.nmf.prob.peptide <- contacts.pTCR.a %>% 
  merge(markup.seq %>% select(pdb.id, peptide.len)) %>%  
  merge(mhc_classes) %>% 
  group_by(pdb.id, residue.index.to, peptide.len, mhc.class) %>% 
  summarise(n.cont = n()) %>% 
  group_by(peptide.len, mhc.class) %>% 
  mutate(n.pdb = n_distinct(pdb.id)) %>% 
  ungroup() %>% 
  group_by(mhc.class, peptide.len, residue.index.to) %>% 
  summarise(freq = sum(n.cont) / n.pdb) %>% 
  distinct()

rs.nmf.prob.tcr <- contacts.pTCR.a %>%
  merge(cdr3_seq %>% select(-region.sequence)) %>% 
  filter(region.type.from == "CDR3") %>% 
  group_by(pdb.id, chain.type.from, pos.from, cdr3.len) %>% 
  summarise(n.cont = n()) %>% 
  group_by(chain.type.from, cdr3.len) %>% 
  mutate(n.pdb = n_distinct(pdb.id)) %>% 
  ungroup() %>% 
  group_by(chain.type.from, pos.from, cdr3.len) %>% 
  summarise(freq = sum(n.cont) / n.pdb) %>% 
  distinct()

#contacts.pTCR.a %>%
#  merge(cdr3_seq %>% select(-region.sequence)) %>% 
#  filter(region.type.from == "CDR3") %>% 
#  filter(pos.from == cdr3.len)
```

```{r}
rs.nmf.peptide <- contacts.pTCR.a %>% 
  merge(markup.seq %>% select(pdb.id, peptide.len)) %>% 
  select(pdb.id, residue.index.to, residue.aa.to, peptide.len) %>% 
  merge(rs.nmf.prob.peptide) %>% 
  #mutate(total = sum(1 * freq)) %>% 
  group_by(residue.aa.to) %>% 
  #summarise(total.to = sum(1 * freq) / total) %>% 
  summarise(total.to = sum(1 * freq)) %>%
  distinct() 

rs.nmf.tcr <- contacts.pTCR.a %>% 
  filter(region.type.from == "CDR3") %>% 
  merge(cdr3_seq %>% select(-region.sequence)) %>%
  select(pdb.id, chain.type.from, pos.from, residue.aa.from, cdr3.len) %>% 
  merge(rs.nmf.prob.tcr) %>% 
  #mutate(total = sum(1 * freq)) %>% 
  group_by(residue.aa.from) %>% 
  #summarise(total.from = sum(1 * freq) / total) %>% 
  summarise(total.from = sum(1 * freq)) %>%
  distinct() 

rs.nmf.a <- rs.nmf.peptide %>% rename(residue.aa = residue.aa.to) %>% 
  merge(rs.nmf.tcr %>% rename(residue.aa = residue.aa.from), all = T) %>% 
  mutate(total.from = replace_na(total.from, 0)) %>% 
  mutate(potential = "TCRen.a.nmf")

rs.nmf.s <- rs.nmf.a %>% 
  mutate(total = total.to + total.from) %>% 
  select(residue.aa, total.to = total, total.from = total) %>% 
  mutate(potential = "TCRen.s.nmf")

rs.nmf <- rs.nmf.a %>% 
  rbind(rs.nmf.s) %>% 
  group_by(potential) %>% 
  mutate(freq.from = total.from / sum(total.from),
         freq.to = total.to / sum(total.to)) %>% 
  ungroup %>% 
  select(residue.aa, potential, freq.from, freq.to)

```

## TCRen calculation with different reference states

```{r}
prepare_rs <- function(rs) {
  expand.grid(residue.aa.from = aa.levels %>% as.character(),
              residue.aa.to = aa.levels %>% as.character()) %>% 
    merge(rs %>% select(residue.aa.from = residue.aa, freq.from, potential), all = T) %>% 
    merge(rs %>% select(residue.aa.to = residue.aa, freq.to, potential), all = T)
}

rs.cf.p <- prepare_rs(rs.cf)
rs.mf.p <- prepare_rs(rs.mf)
rs.nmf.p <- prepare_rs(rs.nmf)


counts.rs <- rbind(
  rbind(
    counts.asym.p %>% mutate(potential = "TCRen.a.cf"),
    counts.sym.p %>% mutate(potential = "TCRen.s.cf")
  ) %>% merge(rs.cf.p),
  rbind(
    counts.asym.p %>% mutate(potential = "TCRen.a.mf"),
    counts.sym.p %>% mutate(potential = "TCRen.s.mf")
  ) %>% merge(rs.mf.p, all = T),
  rbind(
    counts.asym.p %>% mutate(potential = "TCRen.a.nmf"),
    counts.sym.p %>% mutate(potential = "TCRen.s.nmf")
  ) %>% merge(rs.nmf.p, all = T) 
)
  
TCRen <- counts.rs %>%   
  mutate(freq.exp = freq.from * freq.to,
         odds = log(freq.obs / freq.exp),
         value = -odds) %>% 
  filter(residue.aa.from != "C") %>%  # C is never occured in CDR3 regions according to TCR repertoire sequencing
  select(residue.aa.from, residue.aa.to, potential, value)
```


# Potentials from AAindex

```{r, fig.width=8, fig.height=8}
aaindex_list <- c("MOOG990101","KESO980101","MIYS990106","BRYS930101", "MICC010101","THOP960101","TOBD000102") #,"VENM980101"

plot_aaindex_pot_cor <- function(aaindex) {
  aaindex %>% 
  .[,-c(2:6)] %>% 
  column_to_rownames("AAIndexID") %>% 
  t() %>% 
  cor() 
}

plot_aaindex_pot_cor(fread("aaindex_potentials.csv")) %>% 
  corrplot()
plot_aaindex_pot_cor(fread("aaindex_potentials.csv") %>% filter(AAIndexID %in% aaindex_list)) %>% 
  corrplot.mixed(tl.cex = .75)
```

```{r}
aaindex.a <- fread("aaindex_potentials.csv") %>% 
  .[,-c(2:6)] %>% 
  melt(id = c("AAIndexID"), variable.name = "contact_pair") %>% 
  separate(contact_pair, into = c("residue.aa.from", "residue.aa.to"), sep = 1:2) %>% 
  rename(potential = AAIndexID) 

aaindex.pot <- aaindex.a %>% 
  rbind(aaindex.a %>% rename(residue.aa.from = residue.aa.to, residue.aa.to = residue.aa.from)) %>%
  distinct()

aaindex.pot.sel <- aaindex.pot %>% 
  filter(potential %in% aaindex_list) %>% 
  mutate(potential = substr(potential, 1, 4))
```

```{r}
library(hablar)
aaindex_list

aaindex.pot.sel %>% select(residue.aa.to, residue.aa.from, potential, value) %>% 
  rbind(TCRen %>% select(residue.aa.to, residue.aa.from, potential, value)) %>% 
  mutate(potential = factor(potential, levels = c("TCRen.a.cf", "TCRen.a.mf","TCRen.a.nmf","TCRen.s.cf","TCRen.s.mf","TCRen.s.nmf", "MOOG","KESO","MIYS","BRYS","MICC","THOP","TOBD"))) %>% 
  rationalize() %>% 
  mutate(pair = paste0(residue.aa.to, residue.aa.from)) %>% 
  select(pair, potential, value) %>%
  mutate(value = as.numeric(replace_na(value, 0))) %>% 
  dcast(pair ~ potential) %>% 
  drop_na() %>% 
  column_to_rownames("pair") %>% 
  cor() %>% corrplot()
```





# PDB shuffling benchmark

```{r}
bench.shuffle <- expand.grid(pdb.id.1 = pdb_nonred,
            pdb.id.2 = pdb_nonred) %>% 
  merge(markup.seq %>% 
          select(pdb.id.1 = pdb.id, 
                 cdr3a.len.1 = cdr3a.len, cdr3b.len.1 = cdr3b.len, peptide.len.1 = peptide.len)) %>% 
  merge(markup.seq %>% 
          select(pdb.id.2 = pdb.id, 
                 cdr3a.len.2 = cdr3a.len, cdr3b.len.2 = cdr3b.len, peptide.len.2 = peptide.len)) %>% 
  filter(cdr3a.len.1 == cdr3a.len.2, cdr3b.len.1 == cdr3b.len.2, peptide.len.1 == peptide.len.2) %>% 
  mutate(real = ifelse(pdb.id.1 == pdb.id.2, T, F)) %>% 
  select(pdb.id.1, pdb.id.2, real)
```

```{r}
bench.shuffle.contacts <- contacts.pTCR.a %>% 
  filter(region.type.from == "CDR3") %>% 
  select(pdb.id.1 = pdb.id, chain.type.from, residue.index.to, residue.aa.to, pos.from) %>% 
  merge(bench.shuffle) %>% 
  merge(resmarkup.a %>% 
          filter(region.type == "CDR3") %>% 
          merge(markup) %>% 
          mutate(pos.from = residue.index - region.start) %>% 
          select(pdb.id.2 = pdb.id, chain.type.from = chain.type, pos.from, residue.aa.from = residue.aa))

bench.shuffle.energy <- bench.shuffle.contacts %>% 
  merge(TCRen) %>% 
  group_by(pdb.id.1, pdb.id.2, real, potential) %>% 
  summarise(value.s = sum(value))

bench.shuffle.energy.chain <- bench.shuffle.contacts %>% 
  merge(TCRen) %>% 
  group_by(pdb.id.1, pdb.id.2, real, chain.type.from, potential) %>% 
  summarise(value.s = sum(value)) %>% 
  rbind(bench.shuffle.energy %>% mutate(chain.type.from = "TRA+TRB")) %>% 
  mutate(chain.type.from = factor(chain.type.from, levels = c("TRA+TRB", "TRA", "TRB")))

```

```{r}
library(plotROC)
library(PRROC)
library(pROC)

theme_roc <- list( 
    geom_vline(xintercept = 1, size = 1),
    geom_abline(intercept = 1, slope = 0, size = 1),
    geom_roc(n.cuts = 0, size = .5),
    style_roc(major.breaks = c(0, 0.25, 0.5, 0.75, 1), minor.breaks = F, theme = theme_classic),
    geom_abline(intercept = 0, slope = 1, linetype="dashed"),
    labs(#color = "Potential : AUC", 
         color = "",
         x = 'FPR', y = 'TPR'),
    scale_x_continuous(expand = c(0, 0)), scale_y_continuous(expand = c(0, 0)),
    theme(aspect.ratio = 1, 
          legend.position = c(.8,.25),
          legend.background = element_blank(),
          legend.text=element_text(size=9),
          legend.key.size = unit(.7, 'lines'),
          )
)
```

```{r}
auc_TCRen <- bench.shuffle.energy %>% 
  group_by(potential) %>% 
  summarise(auc = round(auc(roc(real, value.s, direction = ">", levels = c(F,T))), 2)) %>% 
  mutate(label = paste0(potential, " : ", auc))

pfig3a <- bench.shuffle.energy %>% 
  ggplot(aes(d = real, m = -value.s, color = potential)) +
  theme_roc +
  scale_color_discrete(labels = auc_TCRen$label)

pfig3a
```

```{r}
aaindex.levels <- c("TCRen.a.cf","MOOG","KESO","MIYS","BRYS","MICC","THOP","TOBD") 

bench.shuffle.energy.aaindex <- bench.shuffle.contacts %>% 
  merge(TCRen %>% 
          filter(potential == "TCRen.a.cf") %>% 
          rbind(aaindex.pot.sel)) %>% 
  mutate(potential = factor(potential, levels = aaindex.levels)) %>% 
  group_by(pdb.id.1, pdb.id.2, real, potential) %>% 
  summarise(value.s = sum(value))

auc_aaindex <- bench.shuffle.energy.aaindex %>% 
  group_by(potential) %>% 
  summarise(auc = round(auc(roc(real, value.s, direction = ">", levels = c(F,T))), 2)) %>% 
  mutate(label = paste0(potential, " : ", auc))

pfig3b <- bench.shuffle.energy.aaindex %>% 
  ggplot(aes(d = real, m = -value.s, color = potential)) +
  theme_roc +
  scale_color_discrete(labels = auc_aaindex$label) +
  scale_color_manual("ROC AUC", 
                     labels = auc_aaindex %>% 
                       mutate(potential = str_split_fixed(potential, "\\.", 2)[,1],
                              label = paste0(potential, " : ", auc)) %>% 
                       .$label, values = cols) +
  theme(legend.text.align = 1,
        legend.title.align = 1)

pfig3b
```


```{r}
auc_TCRen_chains <- bench.shuffle.energy.chain %>% 
  filter(potential == "TCRen.a.cf") %>% 
  group_by(chain.type.from) %>% 
  summarise(auc = round(auc(roc(real, value.s, direction = ">", levels = c(F,T))), 2)) %>% 
  mutate(label = paste0(chain.type.from, " : ", auc))

pfig3c <- bench.shuffle.energy.chain %>% 
  filter(potential == "TCRen.a.cf") %>%
  ggplot(aes(d = real, m = -value.s, color = chain.type.from)) +
  theme_roc +
  scale_color_discrete(labels = auc_TCRen_chains$label) +
  ggtitle("TCRen") +
  theme(legend.position = c(.78,.15)) +
  scale_color_manual("ROC AUC", 
                     labels = auc_TCRen_chains %>% 
                       mutate(label = paste0(chain.type.from, " : ", auc)) %>% 
                       .$label, values = c("red", "blue", "deepskyblue")) +
  theme(legend.text.align = 1,
        legend.title.align = 1)

pfig3c
```

```{r}
library(ggpubr)

pfig3d <- contacts.pTCR.a %>% 
  merge(TCRen) %>% 
  filter(potential == "TCRen.a.cf") %>% 
  group_by(pdb.id, chain.type.from) %>% 
  summarise(value.s = sum(value)) %>% 
  dcast(pdb.id ~ chain.type.from) %>% 
  ggplot(aes(x = TRA, y = TRB)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_pubclean() +
  ggtitle("TCRen")

pfig3d
```

```{r, fig.width=11, fig.height=3.5}
ggdraw() +
  draw_plot(pfig3b, x = 0, y = 0, width = .33, height = .95) +
  draw_plot(pfig3c, x = .33, y = 0, width = .33, height = 1) +
  draw_plot(pfig3d, x = .66, y = 0, width = .33, height = 1) +
  draw_plot_label(label = c("A", "B", "C"), size = 10,
                  x = c(0, .33, .66), y = c(1, 1, 1))
```

```{r, fig.width=8, fig.height=8}
ggdraw() +
  draw_plot(pfig3a, x = 0, y = .52, width = .5, height = .45) +
  draw_plot(pfig3b, x = .5, y = .52, width = .5, height = .45) +
  draw_plot(pfig3c, x = 0, y = 0, width = .5, height = .48) +
  draw_plot(pfig3d, x = .55, y = 0, width = .45, height = .48) +
  draw_plot_label(label = c("A", "B", "C", "D"), size = 10,
                  x = c(.02, .5, .02, .5), y = c(1, 1, .5, .5))
```

# Cognate/unrelated peptides benchmark

```{r}
library(stringi)

aa.re <- "[LFIMVWYCHAGPTSQNDERK]"

peptide_seq <- general %>% 
  filter(#pdb.id %in% pdb_nonred,
         chain.component == "PEPTIDE") %>% 
  select(pdb.id, peptide.pdb = allele.info, peptide.len = seq.length)

bench.peptides <- peptide_seq %>% 
  slice(rep(1:n(), each = 1000)) %>% 
  rowwise() %>% 
  mutate(peptide = paste0(stri_rand_strings(1, 1, aa.re), substr(peptide.pdb, 2, 2), stri_rand_strings(1, peptide.len-3, aa.re), substr(peptide.pdb, peptide.len, peptide.len))) %>% 
  select(pdb.id, peptide) %>% 
  mutate(real = F) %>% 
  rbind(peptide_seq %>% mutate(real = T) %>% select(pdb.id, peptide = peptide.pdb, real))
```

```{r}
max.pep.len <- max(nchar(bench.peptides$peptide))

bench.peptides.res <- bench.peptides %>% 
  rowwise() %>% 
  filter(pdb.id %in% pdb_nonred) %>% 
  select(pdb.id, peptide, real) %>% 
  separate(peptide, into = as.character(0:(max.pep.len - 1)), 
                     sep = 1:max.pep.len, remove = F) %>% 
            melt(id = colnames(bench.peptides), 
                 variable.name = "residue.index.to", value.name = "residue.aa.to") %>% 
            filter(residue.aa.to != "") %>% 
            mutate(residue.index.to = as.integer(as.character(residue.index.to)))

```

```{r}
bench.peptides.contacts <- contacts.pTCR %>%
  filter(pdb.id %in% pdb_nonred) %>% 
  select(pdb.id, chain.type.from, region.type.from, residue.index.from, residue.index.to, residue.aa.from) %>% 
  merge(bench.peptides.res,
          allow.cartesian=TRUE)

bench.peptides.energy <- bench.peptides.contacts %>% 
  merge.data.frame(TCRen) %>% 
  group_by(pdb.id, peptide, real, potential) %>% 
  summarise(value.s = sum(value))
```

```{r}
bench.peptides.ranks <- bench.peptides.energy %>% 
  group_by(pdb.id, potential) %>% 
  mutate(value.rank = rank(value.s) / n() * 100) %>% 
  filter(real)

bench.peptides.ranks.med <- bench.peptides.ranks %>% 
  filter(pdb.id %in% pdb_nonred[1:(length(pdb_nonred)-10)]) %>%
  group_by(potential) %>% 
  summarise(rank.median = round(median(value.rank), 0))
```

# Select which potential is the best

```{r}
pfig2a <- bench.peptides.ranks %>% 
  filter(pdb.id %in% pdb_nonred) %>% 
  ggplot(aes(x = potential, y = value.rank, fill = potential)) +
  geom_violin() +
  geom_boxplot(color = 'black', alpha = 0, width = .1) +
  #scale_x_discrete(guide = guide_axis(angle = 30)) +
  xlab("") + ylab("Cognate epitope rank") +
  guides(fill=FALSE) +
  geom_text(data = bench.peptides.ranks.med, 
            aes(y = rank.median, label = paste0(rank.median, "%")),
            size = 5, vjust = .3, hjust = -.3) +
  theme_pubclean() 

pfig2a
```

```{r, fig.width=7, fig.height=7}
ggdraw() +
  draw_plot(pfig2a, x = 0, y = .5, width = 1, height = .5) +
  draw_plot(pfig2c + theme(legend.position = c(.74,.23)), 
            x = 0, y = 0, width = .5, height = .5) +
  draw_plot(pfig3a + theme(legend.position = c(.74,.23)) + xlab("FPR") + ylab("TPR"), 
            x = .5, y = 0, width = .5, height = .5) +
  draw_plot_label(label = c("B", "C", "D"), size = 11,
                  x = c(0, 0, .5), y = c(1, .5, .5))
```

## Validation on a holdout set

```{r}
fread("PDB_date.csv") %>% 
  merge(tibble(`PDB ID` = toupper(contacts.pTCR$pdb.id %>% unique())), all.y = T) %>% 
  filter(is.na(`Entry ID`) == T)

fread("PDB_date.csv") %>% 
  merge(tibble(`PDB ID` = toupper(pdb_nonred), all.y = T)) %>% 
  arrange(`Release Date`) %>% 
  filter(`Release Date` > "2020-01-01")
```


```{r}
count_contacts <- function(.pdb_list = pdb_nonred, exclude = NULL) {
  contacts.pTCR %>% 
    filter(pdb.id %in% .pdb_list,
           !pdb.id %in% exclude) %>% 
    group_by(residue.aa.from, residue.aa.to) %>% 
    summarise(count = n()) %>% 
    merge(expand.grid(residue.aa.from = aa.levels,
                 residue.aa.to = aa.levels), all = T) %>% 
    mutate(count = replace_na(count, 0))
}

#n.holdout <- round(length(pdb_nonred) / 4)

#pdb_holdout <- pdb_nonred[(length(pdb_nonred)-n.holdout):length(pdb_nonred)]

pdb_holdout <- fread("PDB_date.csv") %>% 
  merge(tibble(`PDB ID` = toupper(pdb_nonred), all.y = T)) %>% 
  arrange(`Release Date`) %>% 
  filter(`Release Date` > "2019-01-01") %>%
  mutate(pdb.id = tolower(`PDB ID`)) %>% 
  .$pdb.id

TCRen.holdout <- calc_TCRen.a.cf(.pdb_list = pdb_nonred, exclude = pdb_holdout, pseudocount = 5) %>% 
  rename(TCRen.holdout = TCRen)

bench.peptides.holdout <- bench.peptides.contacts %>% 
  filter(pdb.id %in% pdb_holdout) %>% 
  merge(TCRen.holdout) %>% 
  group_by(pdb.id, peptide, real) %>% 
  summarise(TCRen.holdout.s = sum(TCRen.holdout)) 

bench.peptides.holdout.ranks <- bench.peptides.holdout %>% 
  group_by(pdb.id) %>% 
  mutate(rank = rank(TCRen.holdout.s) / n()) %>% 
  filter(real == T) 

bench.peptides.holdout.ranks %>% 
  ggplot(aes(x = "TCRen.a.cf (holdout)", y = rank)) +
  geom_quasirandom(width = .2, size = .5, color = "grey") +
  geom_boxplot(alpha = 0, width = .2) +
  theme_bw() +
  theme(aspect.ratio = 2) +
  xlab("") + ylab("Cognate epitope rank") +
  ggtitle("Holdout set")

bench.peptides.holdout.ranks %>% 
  ungroup %>% 
  summarise(median_rank = median(rank))

pdb_covid <- c("7n1e", "7n1f", "7n6e")

bench.peptides.holdout %>% 
  group_by(pdb.id) %>% 
  mutate(rank = rank(TCRen.holdout.s) / n()) %>% 
  filter(real == T) %>% 
  filter(pdb.id %in% pdb_covid)
```

```{r, fig.height=10, fig.width=10}
holdout_labels <- bench.peptides.holdout.ranks %>% 
  filter(!pdb.id %in% pdb_covid) %>% 
  mutate(label = paste0(toupper(pdb.id), " (", round(rank*100), "%)")) %>% 
  arrange(rank) %>% 
  mutate(label = factor(label, levels = .$label)) %>% 
  select(pdb.id, label)


pfigs4a <- bench.peptides.holdout %>% 
  merge(holdout_labels) %>% 
  mutate(label = factor(label, levels = holdout_labels$label)) %>% 
  ggplot(aes(x = "", y = TCRen.holdout.s)) +
  geom_quasirandom(size = .1, color = "grey") +
  geom_boxplot(alpha = 0) +
  geom_point(data = bench.peptides.holdout %>% 
               filter(pdb.id %in% holdout_labels$pdb.id, real) %>% 
               merge(holdout_labels), color = "red") +
  facet_wrap(~label, ncol = 8) +
  theme_bw() +
  #ylim(c(-5, 15)) +
  xlab("") + ylab("TCRen")

pfigs4a
```

```{r}
bench.peptides.covid <- bench.peptides.holdout %>% 
  filter(pdb.id %in% pdb_covid) %>% 
  merge(markup.seq %>% select(pdb.id, peptide.pdb = peptide)) %>% 
  mutate(label1 = toupper(pdb.id), 
         label2 = paste0("\nSARS-CoV2, ", peptide.pdb))

bench.7n6e.cdr3 <- bench.peptides.contacts %>% 
  filter(pdb.id == "7n6e", 
         region.type.from == "CDR3") %>% 
  merge(TCRen.holdout) %>% 
  group_by(pdb.id, peptide, real) %>% 
  summarise(TCRen.holdout.s = sum(TCRen.holdout)) %>% 
  ungroup %>% 
  #mutate(rank = rank(TCRen.holdout.s)) %>% 
  #filter(real) %>% 
  mutate(label1 = "7N6E, CDR3 only",
         label2 = "\nSARS-CoV2 YLQPRTFLL")

bench.peptides.covid.2 <- bench.peptides.covid %>% 
  select(-peptide.pdb) %>% 
  rbind(bench.7n6e.cdr3) %>% 
  group_by(pdb.id, label1, label2) %>% 
  mutate(rank = rank(TCRen.holdout.s)/ n()) %>%  
  filter(real) %>% 
  mutate(label = paste0(label1, " (", round(rank*100), "%)", label2))

pfigs4b <- bench.peptides.covid %>% 
  select(-peptide.pdb) %>% 
  rbind(bench.7n6e.cdr3) %>% 
  merge(bench.peptides.covid.2 %>% select(pdb.id, label1, label2, label)) %>% 
  ggplot(aes(x = "", y = TCRen.holdout.s)) +
  geom_quasirandom(size = .5, color = "grey") +
  geom_boxplot(alpha = 0) +
  geom_point(data = bench.peptides.covid.2,
             color = "red", size = 2) +
  facet_wrap(~label, ncol = 9) +
  theme_bw() +
  ylim(c(-5, 15)) +
  xlab("") + ylab("TCRen")

pfigs4b
```

```{r, fig.width=10, fig.height=12}
ggdraw() +
  draw_plot(pfigs4a, x = 0, y = .3, width = 1, height = .7) +
  draw_plot(pfigs4b, x = 0, y = 0, width = 1, height = .3) +
  draw_plot_label(label = c("A", "B"), size = 10,
                  x = c(0, 0), y = c(1, .3))
```

```{r}
bench.peptides.holdout.ranks %>% 
  rename(peptide.pdb = peptide) %>% 
  merge(vdjdb %>% 
          select(peptide.pdb = Epitope, Epitope.gene, Epitope.species) %>% 
          distinct(peptide.pdb, .keep_all = T) %>% 
          filter(Epitope.gene != ""), all.x = T) %>% 
  merge(iedb %>% 
          select(peptide.pdb = Epitope.2, Epitope.6, Epitope.10) %>% 
          distinct(), all.x = T) %>% 
  arrange(rank)
```


```{r}
contacts_pdb <- function(.pdb.id) {
  contacts.pTCR %>% 
    filter(pdb.id %in% c(.pdb.id)) %>% 
    merge(TCRen.holdout) %>% 
    merge(count_contacts(exclude = pdb_holdout)) %>% 
    mutate(pos.from = paste0(residue.index.from, " (", region.type.from, ")")) %>% 
    select(`PDB ID` = pdb.id, `TCR aa` = residue.aa.from, `peptide aa` = residue.aa.to, 
           `TCR residue` = pos.from, `peptide residue` = residue.index.to,
           TCRen = TCRen.holdout, `count in database` = count) %>% 
    arrange(-TCRen)
}

q <- rbind(
  contacts_pdb("7n6e"),
  contacts_pdb("6vrn"),
  contacts_pdb("6py2"),
  contacts_pdb("6rpa")
)
      
```

```{r}
bench.peptides.contacts %>% 
  filter(pdb.id == "7n6e", 
         region.type.from == "CDR3") %>% 
  merge(TCRen.holdout) %>% 
  group_by(pdb.id, peptide, real) %>% 
  summarise(TCRen.holdout.s = sum(TCRen.holdout)) %>% 
  mutate(rank = rank(TCRen.holdout.s)) %>% 
  ungroup() %>% 
  filter(real)
```


```{r, fig.width=9}
pfigs4 <- ggdraw() +
  draw_plot(pfigs4a, x = 0, y = .5, width = .2, height = .5) +
  draw_plot(pfigs4b, x = .25, y = 0, width = .75, height = 1) +
  draw_plot(pfigs4c, x = 0, y = 0, width = .2, height = .5) +
  draw_plot_label(label = c("A", "B", "C"), size = 10,
                  x = c(0, .25, 0), y = c(1, 1, .5))

pfigs4
```


## Performance comparison with AAindex pairwise potentials

```{r}
bench.peptides.aaindex <- bench.peptides.contacts %>% 
  merge(TCRen %>% 
          filter(potential == "TCRen.a.cf") %>% 
          rbind(aaindex.pot.sel)) %>% 
  group_by(pdb.id, peptide, real, potential) %>% 
  summarise(value.s = sum(value))
```

```{r}
bench.peptides.aaindex.ranks <- bench.peptides.aaindex %>% 
  mutate(potential = factor(potential, levels = aaindex.levels)) %>% 
  group_by(pdb.id, potential) %>% 
  mutate(value.rank = rank(value.s) / n() * 100) %>% 
  filter(real)

bench.peptides.aaindex.ranks.med <- bench.peptides.aaindex.ranks %>% 
  filter(pdb.id %in% pdb_nonred[1:(length(pdb_nonred)-10)]) %>%
  group_by(potential) %>% 
  summarise(rank.median = round(median(value.rank), 0))
```

```{r}
pfig2b <- bench.peptides.aaindex.ranks %>% 
  filter(pdb.id %in% pdb_nonred) %>% 
  ggplot(aes(x = potential, y = value.rank, fill = potential)) +
  geom_violin() +
  geom_boxplot(color = 'black', alpha = 0, width = .1) +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  xlab("") + ylab("Cognate epitope rank") +
  guides(fill=FALSE) +
  geom_text(data = bench.peptides.aaindex.ranks.med, 
            aes(y = rank.median, label = paste0(rank.median, "%")),
            size = 5, vjust = .3, hjust = -.3) +
  theme_pubclean()

pfig2b
```

```{r}
auc_pep_TCRen <- bench.peptides.energy %>% 
  group_by(potential) %>% 
  summarise(auc = round(auc(roc(real, value.s, direction = ">", levels = c(F,T))), 2)) %>% 
  mutate(label = paste0(potential, " : ", auc))

pfig2c <- bench.peptides.energy %>% 
  ggplot(aes(d = real, m = -value.s, color = potential)) +
  theme_roc +
  scale_color_discrete(labels = auc_pep_TCRen$label)

pfig2c
```

```{r}
auc_pep_aaindex <- bench.peptides.aaindex %>% 
  mutate(potential = factor(potential, levels = aaindex.levels)) %>% 
  group_by(potential) %>% 
  summarise(auc = round(auc(roc(real, value.s, direction = ">", levels = c(F,T))), 2)) %>% 
  mutate(label = paste0(potential, " : ", auc))

pfig2d <- bench.peptides.aaindex %>% 
  mutate(potential = factor(potential, levels = aaindex.levels)) %>% 
  ggplot(aes(d = real, m = -value.s, color = potential)) +
  theme_roc +
  scale_color_discrete(labels = auc_pep_aaindex$label)

pfig2d
```


```{r, fig.width=8, fig.height=12}
pfig2 <- ggdraw() +
  draw_plot(pfig2a, x = 0, y = .66, width = 1, height = .33) +
  draw_plot(pfig2b, x = 0, y = .33, width = 1, height = .33) +
  draw_plot(pfig2c, x = 0, y = 0, width = .5, height = .33) +
  draw_plot(pfig2d, x = .5, y = 0, width = .5, height = .33) +
  draw_plot_label(label = c("A", "B", "C", "D"), size = 10,
                  x = c(0, 0, 0, .5), y = c(1, .66, .33, .33))

pfig2
```

# Prediction of neo-epitope recognized by TCR 302TIL

```{r}
run_mir(pdb_dir = "neoepitope/struct/", arg = "annotate-structures")
run_mir(pdb_dir = "neoepitope/struct/", arg = "compute-pdb-contacts")
```


```{r}
contacts.302TIL <- fread("~/shared-with-me/struct2020/benchmark/final_version/neoepitope/mir_output/atomdist.txt") %>% 
  select(-atom.from, -atom.to, -dist) %>% 
  rbind(fread("~/shared-with-me/struct2020/benchmark/final_version/neoepitope/mir_output/atomdist.txt") %>% 
          select(pdb.id, chain.id.from = chain.id.to, chain.id.to = chain.id.from,
                 residue.index.from = residue.index.to, residue.index.to = residue.index.from)) %>% 
  merge(fread("~/shared-with-me/struct2020/benchmark/final_version/neoepitope/mir_output/general.txt") %>% 
          select(pdb.id, chain.id.from = chain.id, chain.type.from = chain.type)) %>% 
  merge(fread("~/shared-with-me/struct2020/benchmark/final_version/neoepitope/mir_output/general.txt") %>% 
          select(pdb.id, chain.id.to = chain.id, chain.type.to = chain.type)) %>% 
  merge(fread("~/shared-with-me/struct2020/benchmark/final_version/neoepitope/mir_output/resmarkup.txt", blank.lines.skip = T) %>% 
          select(pdb.id, chain.id.from = chain.id, residue.index.from = residue.index,
                 residue.aa.from = residue.aa)) %>% 
  merge(fread("~/shared-with-me/struct2020/benchmark/final_version/neoepitope/mir_output/resmarkup.txt", blank.lines.skip = T) %>% 
          select(pdb.id, chain.id.to = chain.id, residue.index.to = residue.index,
                 residue.aa.to = residue.aa)) %>%
  #filter(pdb.id == "6uk4_TCRpMHCmodels_polyV.pdb") %>% 
  distinct() %>% 
  filter(chain.type.from %in% c("TRA", "TRB"),
         chain.type.to == "PEPTIDE") %>% 
  mutate(type = ifelse(pdb.id == "6uk4.pdb", "PDB",
                         ifelse(pdb.id == "6uk4_TCRpMHCmodels_polyV.pdb", "model", pdb.id)))

neoepitope_real <- "KQWLVWLFL"

candidates <- fread("~/shared-with-me/struct2020/benchmark/final_version/neoepitope/candidates_A0206.tsv") %>% 
  mutate(real = peptide == neoepitope_real) %>% 
  separate(peptide, into = as.character(0:8), 
                     sep = 1:9, remove = F) %>% 
  melt(id = c("peptide", "real"), 
       variable.name = "residue.index.to", value.name = "residue.aa.to") %>%
  mutate(residue.index.to = as.integer(as.character(residue.index.to)))

energy.302TIL <- contacts.302TIL %>%
  select(pdb.id, type, chain.id.from, residue.index.from, residue.index.to, residue.aa.from) %>% 
  merge(candidates,
          allow.cartesian=TRUE) %>% 
  merge.data.frame(TCRen %>% 
                     rbind(calc_TCRen.a.cf(pdb_nonred[pdb_nonred != "6p64"]) %>% 
                             mutate(potential = "LOO.TCRen") %>% 
                             rename(value = TCRen))) %>% 
  #merge(tcren.mean.flat %>% 
  #        mutate(potential = "TCRen.005") %>% 
  #        select(residue.aa.from, residue.aa.to, potential, value = TCRen.005)) %>% 
  group_by(pdb.id, type, peptide, real, potential) %>% 
  summarise(value.s = sum(value))

energy.302TIL %>% 
  ggplot(aes(x = type, y = value.s, color = real)) +
  geom_quasirandom(width = .1) +
  xlab("") + ylab("TCRen") +
  #geom_text_repel(aes(label = ifelse(real == T, "cognate neo-epitope", "")), nudge_x = 1) +
  theme_pubclean() +
  scale_color_manual(values = c("black", "red")) +
  theme(legend.position = "none") +
  theme(axis.ticks = element_blank()) +
  facet_wrap(~potential, scales = "free")

```

```{r}
energy.302TIL %>% 
  filter(potential == "LOO.TCRen", type == "model") %>% 
  ggplot(aes(x = "", y = value.s, color = real)) +
  geom_quasirandom(width = .1) +
  xlab("") + ylab("TCRen") +
  #geom_text_repel(aes(label = ifelse(real == T, "cognate neo-epitope", "")), nudge_x = 1) +
  geom_text_repel(aes(label = ifelse(real == T, "cognate \nneo-epitope", "")), color = "red", box.padding = 0.5, max.overlaps = Inf) +
  theme_pubclean() +
  scale_color_manual(values = c("black", "red")) +
  theme(legend.position = "none") +
  theme(axis.ticks = element_blank()) +
  theme(aspect = 1.6)
```

```{r}
energy.302TIL %>% 
  filter(potential == "LOO.TCRen", type == "model") %>% 
  mutate(value.s = round(value.s, 3)) %>% 
  arrange(value.s)
```


```{r}
resmarkup.302TIL <- fread("~/shared-with-me/struct2020/benchmark/final_version/neoepitope/mir_output/resmarkup.txt", blank.lines.skip = T) %>% 
  merge(fread("~/shared-with-me/struct2020/benchmark/final_version/neoepitope/mir_output/general.txt")) %>% 
  merge(fread("~/shared-with-me/struct2020/benchmark/final_version/neoepitope/mir_output/markup.txt")) %>% 
  mutate(pos = residue.index - region.start)

contacts.302TIL.pos <- contacts.302TIL %>% 
  merge(resmarkup.302TIL %>% 
          select(pdb.id, chain.id.from = chain.id, region.type.from = region.type,
                 residue.index.from = residue.index, pos.from = pos)) %>% 
  merge(resmarkup.302TIL %>% 
          select(pdb.id, chain.id.to = chain.id, region.type.to = region.type, 
                 residue.index.to = residue.index, pos.to = pos))
  
contacts.302TIL.pos %>% 
  group_by(pdb.id) %>% 
  summarise(n = n())
```
### Compare contacts in PDB and in the model
```{r}
contacts.302TIL.comp <- contacts.302TIL.pos %>% 
  mutate(contact = paste0(region.type.from, substr(chain.type.from, 3, 3), "_", pos.from, "_", pos.to)) %>% 
  merge(resmarkup.302TIL %>% 
          filter(pdb.id == "6uk4.pdb", chain.type == "PEPTIDE") %>% 
          select(residue.index.to = residue.index, residue.aa.to.PDB = residue.aa)) %>% 
  select(type, contact, residue.aa.to.PDB, residue.aa.from) %>%
  mutate(in.model = type == "model",
         in.PDB = type == "PDB") %>% 
  group_by(contact, residue.aa.to.PDB, residue.aa.from) %>% 
  summarise(in.model.s = sum(in.model),
            in.PDB.s = sum(in.PDB)) %>% 
  rowwise() %>% 
  mutate(in.both = (in.model.s == 1 && in.PDB.s == 1)) 

contacts.302TIL.comp %>% 
  ungroup %>% 
  select(in.model.s, in.PDB.s, in.both) %>% 
  colSums()

contacts.302TIL.comp %>% 
  rename(residue.aa.to = residue.aa.to.PDB) %>% 
  merge(calc_TCRen.a.cf(exclude = "6p64")) %>% 
  arrange(TCRen)
```

# Test systems with yeast display data

## Birnbaum

```{r}
birnbaum.pdb <- tibble(TCR = c("2b4", "226", "5cc7"),
       pdb.id = c("3qib", "3qiu", "4p2r"))

birnbaum <- birnbaum.pdb %>% 
  merge(fread("yeast_display/Birnbaum.tsv") %>% 
          select(peptide, round_5, TCR) %>% 
          group_by(TCR) %>% 
          slice(1:20)) %>% 
  select(pdb.id, peptide)

clusters.merged %>% 
  filter(pdb.id %in% birnbaum$pdb.id) 

clusters.merged %>% 
  filter(cluster.seq %in% c(43,44,72,73,74) | cluster.cont %in% c(39,60,61,62))
```

```{r}
separate_peptide <- function(df) {
  df %>% 
    separate(peptide, into = as.character(0:(max.pep.len - 1)), 
                     sep = 1:max.pep.len, remove = F) %>% 
            melt(id = c(colnames(bench.peptides), c("peptide.pdb")), 
                 variable.name = "residue.index.to", value.name = "residue.aa.to") %>% 
            filter(residue.aa.to != "") %>% 
            mutate(residue.index.to = as.integer(as.character(residue.index.to)))
}
```


```{r}
pdb_birnbaum <- birnbaum$pdb.id

birnbaum.energy <- bench.peptides %>% 
  ungroup() %>% 
  mutate(peptide.pdb = real == T) %>% 
  rbind(birnbaum %>% 
          mutate(real = T,
                 peptide.pdb = F)) %>% 
  filter(pdb.id %in% pdb_birnbaum) %>% 
  select(pdb.id, peptide, real, peptide.pdb) %>% 
  separate_peptide() %>% 
  merge(contacts.pTCR %>% 
          select(pdb.id, chain.type.from, region.type.from, residue.index.from, residue.index.to, residue.aa.from)) %>% 
  merge(TCRen %>% 
          merge(tibble(pdb.id = pdb_birnbaum)) %>% 
          rbind(TCRen.LOO %>% 
                  filter(pdb.id %in% pdb_birnbaum) %>% 
                  mutate(potential = "LOO.TCRen") %>% 
                  select(residue.aa.from, residue.aa.to, potential, value = TCRen.LOO, pdb.id)))
```

```{r}
birnbaum.energy.s <- birnbaum.energy %>%   
  group_by(pdb.id, peptide, real, peptide.pdb, potential) %>% 
  summarise(value.s = sum(value)) %>% 
  group_by(pdb.id, potential) 

birnbaum.energy.s %>% 
  mutate(rank = rank(value.s) / n()) %>% 
  filter(real) %>% 
  ggplot(aes(x = pdb.id, y = rank)) +
  geom_boxplot() +
  geom_beeswarm() +
  facet_wrap(~potential)
  arrange(pdb.id, rank)

```

```{r, fig.width=3, fig.height=3}
auc_birnbaum <- birnbaum.energy.s %>% 
  filter(pdb.id %in% c("3qib", "3qiu", "4p2r")) %>% 
  filter(potential %in% c("LOO.TCRen")) %>% 
  group_by(pdb.id, potential) %>% 
  summarise(auc = round(auc(roc(real, value.s, direction = ">", levels = c(F,T))), 2)) %>% 
  merge(birnbaum.pdb) %>% 
  mutate(label = paste0("AUC = ", auc))

plot_birnbaum <- function(.TCR) {
  .pdb.id <- auc_birnbaum %>% 
    filter(TCR == .TCR) %>% 
    .$pdb.id
  
  .label <- auc_birnbaum %>% 
    filter(TCR == .TCR) %>% 
    .$label
  
  birnbaum.energy.s %>% 
    filter(potential == "LOO.TCRen",
           pdb.id == .pdb.id) %>% 
    ggplot(aes(d = real, m = -value.s, color = pdb.id)) +
    theme_roc +
    scale_color_discrete(labels = .label) 
}

p.2b4.roc <- plot_birnbaum("2b4")
p.226.roc <- plot_birnbaum("226")
p.5cc7.roc <- plot_birnbaum("5cc7")
```

```{r}
library(ggseqlogo)

logo_birnbaum <- function(.TCR) {
  fread("~/shared-with-me/struct2020/benchmark/final_version/yeast_display/Birnbaum.tsv") %>% 
    filter(TCR == .TCR) %>% 
    arrange(-round_5) %>% 
    head(20) %>% 
    .$peptide %>% 
    ggseqlogo() +
    theme(legend.position = "") +
    ggtitle(.TCR) +
    theme(plot.title = element_text(hjust = 0.5))}

p.2b4.logo <- logo_birnbaum("2b4")
p.226.logo <- logo_birnbaum("226")
p.5cc7.logo <- logo_birnbaum("5cc7")

```

```{r, fig.width=9, fig.height=4}
ggdraw() +
  draw_plot(p.2b4.logo, x = 0, y = .6, width = .33, height = .4) +
  draw_plot(p.226.logo, x = .33, y = .6, width = .33, height = .4) +
  draw_plot(p.5cc7.logo, x = .66, y = .6, width = .33, height = .4) +
  draw_plot(p.2b4.roc, x = 0, y = 0, width = .33, height = .65) +
  draw_plot(p.226.roc, x = .33, y = 0, width = .33, height = .65) +
  draw_plot(p.5cc7.roc, x = .66, y = 0, width = .33, height = .65) +
  draw_plot_label(label = c("E"), size = 10,
                  x = c(0), y = c(1))

```
```{r, fig.width=12, fig.height=12}
ggdraw() +
  draw_plot(pfig2a, x = 0, y = .68, width = .66, height = .32) +
  draw_plot(pfig2c, x = .66, y = .68, width = .33, height = .32) +
  draw_plot(pfig2b, x = 0, y = .36, width = .66, height = .32) +
  draw_plot(pfig2d, x = .66, y = .36, width = .33, height = .32) +
  draw_plot(p.2b4.logo, x = 0, y = .22, width = .33, height = .12) +
  draw_plot(p.226.logo, x = .33, y = .22, width = .33, height = .12) +
  draw_plot(p.5cc7.logo, x = .66, y = .22, width = .33, height = .12) +
  draw_plot(p.2b4.roc, x = 0, y = 0, width = .33, height = .23) +
  draw_plot(p.226.roc, x = .33, y = 0, width = .33, height = .23) +
  draw_plot(p.5cc7.roc, x = .66, y = 0, width = .33, height = .23) +
  draw_plot_label(label = c("A", "B", "C", "D", "E"), size = 12,
                  x = c(0, 0, .66, .66, 0), y = c(1, .68, 1, .68, .35))

```


# MD

```{r}
md.mir.cont <- md.mir.a %>% 
  group_by(pdb.id, peptide, motif, replica, chain.id.from, chain.id.to, 
           residue.index.from, residue.index.to, residue.aa.from, residue.aa.to) %>% 
  mutate(frame = as.numeric(frame)) %>% 
  summarise(freq = n() / (max(frame) - min(frame) + 1))
```

```{r}
md.mir.cont.e <- md.mir.cont  %>% 
  filter((pdb.id == "6am5_al2.pdb" & motif %in% c("GIG", "no_motif")) | 
           (pdb.id == "6amu_al2.pdb" & motif %in% c("DRG", "no_motif"))) %>% 
  merge(calc_TCRen.a.cf.2(.pdb_list = pdb_nonred, pseudocount = 2) %>% 
          mutate(potential = "TCRen") %>% 
          rename(value = TCRen)) %>% 
  group_by(pdb.id, peptide, motif, replica, potential) %>% 
  summarise(value.replica = sum(value)) %>%  
  group_by(pdb.id, peptide, motif, potential) %>% 
  summarise(value.s = mean(value.replica)) 

md.mir.cont.e %>% 
  ggplot(aes(x = motif, y = value.s)) +
  facet_wrap(~ pdb.id + potential, scales = 'free') +
  geom_quasirandom(size = .1) +
  geom_boxplot(alpha = 0) +
  theme_bw()

 
stat.test <- md.mir.cont.e %>%
    group_by(pdb.id) %>%
    wilcox_test(value.s ~ motif) %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance()

#stat.test <- stat.test %>% add_y_position()
stat.test
```

```{r}
TCRen.2 <- #calc_TCRen.a.cf(.pdb_list = pdb_nonred, pseudocount = .05) %>% 
  calc_TCRen.a.cf(.pdb_list = pdb_nonred, exclude = pdb_holdout, pseudocount = 2) %>% 
          mutate(potential = "TCRen") %>% 
          rename(value = TCRen)

md.mir.cont.e <- md.mir.cont  %>% 
  filter((pdb.id == "6am5_al2.pdb" & motif %in% c("GIG", "no_motif")) | 
           (pdb.id == "6amu_al2.pdb" & motif %in% c("DRG", "no_motif"))) %>% 
  merge(TCRen.2) %>% 
  group_by(pdb.id, peptide, motif, replica, potential) %>% 
  summarise(value.replica = sum(value)) %>%  
  group_by(pdb.id, peptide, motif, potential) %>% 
  summarise(value.s = mean(value.replica)) 

md.mir.cont.e %>% 
  ggplot(aes(x = motif, y = value.s)) +
  facet_wrap(~ pdb.id + potential, scales = 'free') +
  geom_quasirandom(size = .1) +
  geom_boxplot(alpha = 0) +
  theme_bw()

 
stat.test <- md.mir.cont.e %>%
    group_by(pdb.id) %>%
    wilcox_test(value.s ~ motif) %>%
    adjust_pvalue(method = "bonferroni") %>%
    add_significance()

#stat.test <- stat.test %>% add_y_position()
stat.test
```

```{r}
md.mir.frame0 <- md.mir.a  %>% 
  filter(frame == 0) %>% 
  merge(calc_TCRen.a.cf.2(.pdb_list = pdb_nonred, pseudocount = 2) %>% 
          mutate(potential = "TCRen") %>% 
          rename(value = TCRen)) %>% 
  group_by(pdb.id, peptide, motif, replica, potential) %>% 
  summarise(value.replica = sum(value)) %>%  
  group_by(pdb.id, peptide, motif, potential) %>% 
  summarise(value.s = mean(value.replica),
            n.replica = n())

md.mir.cont.e %>% 
  rename(MD = value.s) %>% 
  merge(md.mir.frame0 %>% rename(frame0 = value.s)) %>% 
  ggplot(aes(x = frame0, y = MD)) +
  geom_point() +
  geom_smooth(method = "lm")

md.mir.a

```

# Amino acid composition of IEDB epitopes

```{r}
iedb <- read.csv("~/MHC_epitopes/mhc_ligand_full.csv")
```

```{r}
iedb %>% 
  filter(Epitope.1 == "Linear peptide") %>% 
  select()
```

```{r, fig.width=10, fig.height=5}
iedb %>% 
  filter(Epitope.1 == "Linear peptide",
         Assay.5 == "Positive") %>% 
  select(Epitope.2, MHC) %>% 
  distinct() %>% 
  mutate(pep.len = nchar(Epitope.2)) %>% 
  filter(pep.len < 15) %>% 
  group_by(MHC) %>% 
  #summarise(n.lig = n()) %>% 
  #filter(n.lig > 1000) %>% 
  mutate(n.lig = n()) %>%  
  filter(n.lig > 1000) %>% 
  select(-n.lig) %>% 
  filter(substr(MHC, 1, 6) %in% c("HLA-A*", "HLA-B*", "HLA-C*")) %>% 
  separate(Epitope.2, into = paste0("p", 1:15), sep = 1:15, remove = F) %>% 
  select(-pep.len) %>% 
  melt(id = c("Epitope.2", "MHC")) %>% 
  group_by(MHC, value) %>% 
  summarise(n = n()) %>% 
  filter(value != "") %>% 
  mutate(fraction = n / sum(n)) %>% 
  ggplot(aes(x = factor(value, levels = aa.levels), y = fraction)) +
  geom_boxplot() +
  geom_line(data = aa.freq.interface, aes(x = factor(aminoacid, levels = aa.levels), y = freq, color = type, group = type)) +
  theme_pubclean() +
  xlab("") + ylab("fraction")
  #ggplot(aes(x = factor(value, levels = aa.levels), y = fraction, color = MHC, group = MHC)) +
  #geom_line()
```

```{r, fig.width=10, fig.height=5}
iedb %>% 
  filter(Epitope.1 == "Linear peptide",
         Assay.5 == "Positive") %>% 
  select(Epitope.2, MHC) %>% 
  distinct() %>% 
  mutate(pep.len = nchar(Epitope.2)) %>% 
  filter(pep.len == 9) %>% 
  group_by(MHC) %>% 
  #summarise(n.lig = n()) %>% 
  #filter(n.lig > 1000) %>% 
  mutate(n.lig = n()) %>%  
  filter(n.lig > 1000) %>% 
  select(-n.lig) %>% 
  filter(substr(MHC, 1, 6) %in% c("HLA-A*", "HLA-B*", "HLA-C*")) %>% 
  separate(Epitope.2, into = paste0("p", 1:9), sep = 1:9, remove = F) %>% 
  select(-pep.len) %>% 
  melt(id = c("Epitope.2", "MHC")) %>% 
  filter(variable != "p2", variable != "p9") %>% 
  group_by(MHC, value) %>% 
  summarise(n = n()) %>% 
  filter(value != "") %>% 
  mutate(fraction = n / sum(n)) %>% 
  ggplot(aes(x = factor(value, levels = aa.levels), y = fraction)) +
  geom_boxplot() +
  geom_line(data = aa.freq.interface, aes(x = factor(aminoacid, levels = aa.levels), y = freq, color = type, group = type)) +
  theme_pubclean() +
  xlab("") + ylab("fraction")
  #ggplot(aes(x = factor(value, levels = aa.levels), y = fraction, color = MHC, group = MHC)) +
  #geom_line()
```

